\documentclass{ximera}
\input{../preamble.tex}

\title{Exercises} \license{CC BY-NC-SA 4.0}

\begin{document}

\begin{abstract}
\end{abstract}
\maketitle

\begin{onlineOnly}
\section*{Exercises}
\end{onlineOnly}




\begin{problem}\label{exer:9.3.1}   Find a particular solution.  

$y'''-6y''+11y'-6y=-e^{-x}(4+76x-24x^2)$
\end{problem}

\begin{problem}\label{exer:9.3.2}   Find a particular solution. 

$y'''-2y''-5y'+6y=e^{-3x}(32-23x+6x^2)$
\end{problem}

\begin{problem}\label{exer:9.3.3}   Find a particular solution.  

$4y'''+8y''-y'-2y=-e^x(4+45x+9x^2)$
\end{problem}

\begin{problem}\label{exer:9.3.4}   Find a particular solution. 

$y'''+3y''-y'-3y=e^{-2x}(2-17x+3x^2)$
\end{problem}

\begin{problem}\label{exer:9.3.5}   Find a particular solution. 

$y'''+3y''-y'-3y=e^x(-1+2x+24x^2+16x^3)$
\end{problem}

\begin{problem}\label{exer:9.3.6}   Find a particular solution.  

$y'''+y''-2y=e^x(14+34x+15x^2)$
\end{problem}

\begin{problem}\label{exer:9.3.7}   Find a particular solution.   

$4y'''+8y''-y'-2y=-e^{-2x}(1-15x)$
\end{problem}

\begin{problem}\label{exer:9.3.8}   Find a particular solution.  

$y'''-y''-y'+y=e^x(7+6x)$
\end{problem}

\begin{problem}\label{exer:9.3.9}   Find a particular solution. 

$2y'''-7y''+4y'+4y=e^{2x}(17+30x)$
\end{problem}

\begin{problem}\label{exer:9.3.10}   Find a particular solution.

$y'''-5y''+3y'+9y=2e^{3x}(11-24x^2)$
\end{problem}

\begin{problem}\label{exer:9.3.11}   Find a particular solution.

$y'''-7y''+8y'+16y=2e^{4x}(13+15x)$
\end{problem}

\begin{problem}\label{exer:9.3.12}   Find a particular solution.

$8y'''-12y''+6y'-y=e^{x/2}(1+4x)$
\end{problem}

\begin{problem}\label{exer:9.3.13}   Find a particular solution.   $y^{(4)}+3y'''-3y''-7y'+6y=-e^{-x}(12+8x-8x^2)$
\end{problem}

\begin{problem}\label{exer:9.3.14}   Find a particular solution.   $y^{(4)}+3y'''+y''-3y'-2y=-3e^{2x}(11+12x)$
\end{problem}

\begin{problem}\label{exer:9.3.15}
$y^{(4)}+8y'''+24y''+32y'=-16e^{-2x}(1+x+x^2-x^3)$
\end{problem}

\begin{problem}\label{exer:9.3.16}   Find a particular solution.

$4y^{(4)}-11y''-9y'-2y=-e^x(1-6x)$

\begin{solution}
If $y=ue^x$, then $4y^{(4)}-11y''-9y'-2y=e^x[
4(u^{(4)}+4u'''+6u''+4u'+u) -11(u''+2u'+u) -9(u'+u) -2u]
=e^x(4u^{(4)}+16u'''+13u''-15u'-18u)$. Let $u_p=A+Bx$ where
$-(18A+15B)-18Bx=-1+6x$. Then $B=-1/3$, $A=1/3$, and
$y_p=\frac{e^x}{3}(1-x)$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:9.3.17}   Find a particular solution.

$y^{(4)}-2y'''+3y'-y=e^x(3+4x+x^2)$
\end{problem}

\begin{problem}\label{exer:9.3.18}   Find a particular solution.   $y^{(4)}-4y'''+6y''-4y'+2y=e^{2x}(24+x+x^4)$
\end{problem}

\begin{problem}\label{exer:9.3.19}   Find a particular solution.   $2y^{(4)}+5y'''-5y'-2y=18e^x(5+2x)$
\end{problem}

\begin{problem}\label{exer:9.3.20}   Find a particular solution.

$y^{(4)}+y'''-2y''-6y'-4y=-e^{2x}(4+28x+15x^2)$
\end{problem}

\begin{problem}\label{exer:9.3.21}   Find a particular solution.   
$2y^{(4)}+y'''-2y'-y=3e^{-x/2}(1-6x)$
\end{problem}

\begin{problem}\label{exer:9.3.22}   Find a particular solution.   
$y^{(4)}-5y''+4y=e^x(3+x-3x^2)$
\end{problem}

\begin{problem}\label{exer:9.3.23}  
Find a particular solution.

$y^{(4)}-2y'''-3y''+4y'+4y=e^{2x}(13+33x+18x^2)$
\end{problem}

\begin{problem}\label{exer:9.3.24}  
Find a particular solution.

$y^{(4)}-3y'''+4y'=e^{2x}(15+26x+12x^2)$
\end{problem}

\begin{problem}\label{exer:9.3.25}   Find a particular solution. 

$y^{(4)}-2y'''+2y'-y=e^x(1+x)$
\end{problem}

\begin{problem}\label{exer:9.3.26}   Find a particular solution.   
$2y^{(4)}-5y'''+3y''+y'-y=e^x(11+12x)$
\end{problem}

\begin{problem}\label{exer:9.3.27}   Find a particular solution.

$y^{(4)}+3y'''+3y''+y'=e^{-x}(5-24x+10x^2)$
\end{problem}

\begin{problem}\label{exer:9.3.28}   Find a particular solution.   $y^{(4)}-7y'''+18y''-20y'+8y=e^{2x}(3-8x-5x^2)$
\end{problem}

\begin{problem}\label{exer:9.3.29}   Find a particular solution.

$y'''-y''-4y'+4y=e^{-x}\left[(16+10x)\cos
x+(30-10x)\sin x\right]$
\end{problem}

\begin{problem}\label{exer:9.3.30}   Find a particular solution.

$y'''+y''-4y'-4y=e^{-x}\left[(1-22x)\cos
2x-(1+6x)\sin2x\right]$
\end{problem}

\begin{problem}\label{exer:9.3.31}   Find a particular solution.

$y'''-y''+2y'-2y=e^{2x}[(27+5x-x^2)\cos
x+(2+13x+9x^2)\sin x]$
\end{problem}

\begin{problem}\label{exer:9.3.32}   Find a particular solution.

$y'''-2y''+y'-2y=-e^x[(9-5x+4x^2)\cos
2x-(6-5x-3x^2)\sin2x]$
\end{problem}

\begin{problem}\label{exer:9.3.33}   Find a particular solution.

$y'''+3y''+4y'+12y=8\cos2x-16\sin2x$
\end{problem}

\begin{problem}\label{exer:9.3.34}   Find a particular solution. 

$y'''-y''+2y=e^x[(20+4x)\cos x-(12+12x)\sin
 x]$
\end{problem}
 
\begin{problem}\label{exer:9.3.35}   Find a particular solution.

$y'''-7y''+20y'-24y=-e^{2x}[(13-8x)\cos
2x-(8-4x)\sin2x]$
\end{problem}

\begin{problem}\label{exer:9.3.36}   Find a particular solution. 

$y'''-6y''+18y'=-e^{3x}[(2-3x)\cos
3x-(3+3x)\sin3x]$
\end{problem}

\begin{problem}\label{exer:9.3.37}   Find a particular solution.   $y^{(4)}+2y'''-2y''-8y'-8y=e^x(8\cos
x+16\sin x)$
\end{problem}

\begin{problem}\label{exer:9.3.38}   Find a particular solution.   $y^{(4)}-3y'''+2y''+2y'-4y=e^x(2\cos2x
-\sin2x)$
\end{problem}

\begin{problem}\label{exer:9.3.39}
$y^{(4)}-8y'''+24y''-32y'+15y=e^{2x}(15x\cos2x+32\sin2x)$
\end{problem}

\begin{problem}\label{exer:9.3.40}
$y^{(4)}+6y'''+13y''+12y'+4y=e^{-x}[(4-x)\cos x-(5+x)\sin x]$
\end{problem}

\begin{problem}\label{exer:9.3.41}   Find a particular solution.   $y^{(4)}+3y'''+2y''-2y'-4y=-e^{-x}
(\cos x-\sin x)$
\end{problem}

\begin{problem}\label{exer:9.3.42}   Find a particular solution.   $y^{(4)}-5y'''+13y''-19y'+10y=e^x
(\cos2x+\sin2x)$
\end{problem}

\begin{problem}\label{exer:9.3.43}
$y^{(4)}+8y'''+32y''+64y'+39y=e^{-2x}[(4-15x)\cos3x-(4+15x)\sin
3x]$
\end{problem}

\begin{problem}\label{exer:9.3.44}   Find a particular solution.   $y^{(4)}-5y'''+13y''-19y'+10y=e^x[(7+8x)\cos
2x+(8-4x)\sin2x]$
\end{problem}

\begin{problem}\label{exer:9.3.45}   Find a particular solution.   $y^{(4)}+4y'''+8y''+8y'+4y=-2e^{-x}
(\cos x-2\sin x)$
\end{problem}

\begin{problem}\label{exer:9.3.46}   Find a particular solution.   $y^{(4)}-8y'''+32y''-64y'+64y=e^{2x}
(\cos2x-\sin2x)$

\begin{solution}
If $y=ue^{2x}$, then
$y^{(4)}-8y'''+32y''-64y'+64y+4y=e^{2x}[
(u^{(4)}+8u'''+24u''+32u'+16u) -8(u'''+6u''+12u'+8u) +32(u''+4u'+4u)
-64(u'+2u) +64u]=e^{2x}(u^{(4)}+8u''+16u)$. Since $\cos2x$,  $\sin2x$, $x\cos2x$,
and $x\sin2x$ satisfy $u^{(4)}+8u''+16u=0$,
let $u_p=x^2(A\cos2x+B\sin2x)$ where $-32A=1$ and $-32B=-1$. Then
$A=-1/32$, $B=1/32$, and
 $y_p=-\frac{x^2e^{2x}}{32}(\cos2x-\sin2x)$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:9.3.47}   Find a particular solution.   $y^{(4)}-8y'''+26y''-40y'+25y=e^{2x}[3\cos
x-(1+3x)\sin x]$
\end{problem}

\begin{problem}\label{exer:9.3.48}   Find a particular solution.   $y'''-4y''+5y'-2y=e^{2x}-4e^x-2\cos x+4\sin x$
\end{problem}

\begin{problem}\label{exer:9.3.49}   Find a particular solution.   $y'''-y''+y'-y=5e^{2x}+2e^x-4\cos x+4\sin x$
\end{problem}

\begin{problem}\label{exer:9.3.50}   Find a particular solution.   $y'''-y'=-2(1+x)+4e^x-6e^{-x}+96e^{3x}$
\end{problem}

\begin{problem}\label{exer:9.3.51}   Find a particular solution.   $y'''-4y''+9y'-10y=10e^{2x}+20e^x\sin2x-10$
\end{problem}

\begin{problem}\label{exer:9.3.52}   Find a particular solution.

$y'''+3y''+3y'+y=12e^{-x}+9\cos2x-13\sin2x$
\end{problem}

\begin{problem}\label{exer:9.3.53}   Find a particular solution.

$y'''+y''-y'-y=4e^{-x}(1-6x)-2x\cos x+2(1+x)\sin
x $
\end{problem}

\begin{problem}\label{exer:9.3.54}   Find a particular solution.

$y^{(4)}-5y''+4y=-12e^x+6e^{-x}+10\cos x$
\end{problem}

\begin{problem}\label{exer:9.3.55}   Find a particular solution.

$y^{(4)}-4y'''+11y''-14y'+10y=-e^x(\sin
x+2\cos2x)$
\end{problem}

\begin{problem}\label{exer:9.3.56}   Find a particular solution.   $y^{(4)}+2y'''-3y''-4y'+4y=2e^x(1+x)+e^{-2x}$
\end{problem}

\begin{problem}\label{exer:9.3.57}   Find a particular solution.   $y^{(4)}+4y=\sinh x\cos x-\cosh x\sin x$
\end{problem}

\begin{problem}\label{exer:9.3.58}
$y^{(4)}+5y'''+9y''+7y'+2y=e^{-x}(30+24x)-e^{-2x}$
\end{problem}

\begin{problem}\label{exer:9.3.59}   Find a particular solution.   $y^{(4)}-4y'''+7y''-6y'+2y=e^x(12x-2\cos
x+2\sin x)$
\end{problem}

\begin{problem}\label{exer:9.3.60}   Find the general solution.

$y'''-y''-y'+y=e^{2x}(10+3x)$
\end{problem}

\begin{problem}\label{exer:9.3.61}   Find the general solution. 

$y'''+y''-2y=-e^{3x}(9+67x+17x^2)$
\end{problem}

\begin{problem}\label{exer:9.3.62}   Find the general solution.

$y'''-6y''+11y'-6y=e^{2x}(5-4x-3x^2)$
\end{problem}

\begin{problem}\label{exer:9.3.63}   Find the general solution. 

$y'''+2y''+y'=-2e^{-x}(7-18x+6x^2)$
\end{problem}

\begin{problem}\label{exer:9.3.64}   Find the general solution. 

$y'''-3y''+3y'-y=e^x(1+x)$
\end{problem}

\begin{problem}\label{exer:9.3.65}  Find the general solution.

$y^{(4)}-2y''+y=-e^{-x}(4-9x+3x^2)$
\end{problem}

\begin{problem}\label{exer:9.3.66}   Find the general solution. 

$y'''+2y''-y'-2y=e^{-2x}\left[(23-2x)\cos
x+(8-9x)\sin x\right]$
\end{problem}

\begin{problem}\label{exer:9.3.67}   Find the general solution.

$y^{(4)}-3y'''+4y''-2y'=e^x\left[(28+6x)\cos
2x+(11-12x)\sin2x\right]$
\end{problem}

\begin{problem}\label{exer:9.3.68}   Find the general solution. 

$y^{(4)}-4y'''+14y''-20y'+25y=e^x\left[(2+6x)\cos
2x+3\sin2x\right]$
\end{problem}

\begin{problem}\label{exer:9.3.69} Solve the initial value problem and graph the solution.

$y'''-2y''-5y'+6y=2e^x(1-6x),\quad
y(0)=2,
\quad y'(0)=7,\quad y''(0)=9$
\end{problem}

\begin{problem}\label{exer:9.3.70} Solve the initial value problem and graph the solution.

$y'''-y''-y'+y=-e^{-x}(4-8x),\quad
y(0)=2,
\quad y'(0)=0,\quad y''(0)=0$
\end{problem}

\begin{problem}\label{exer:9.3.71} Solve the initial value problem and graph the solution. 

$4y'''-3y'-y=e^{-x/2}(2-3x),\quad
y(0)=-1,
\quad y'(0)=15,\quad y''(0)=-17$
\end{problem}

\begin{problem}\label{exer:9.3.72}  Solve the initial value problem and graph the solution.

$y^{(4)}+2y'''+2y''+2y'+y=e^{-x}(20-12x),\,
y(0)=3,\; y'(0)=-4,\; y''(0)=7,\; y'''(0)=-22$
\end{problem}

\begin{problem}\label{exer:9.3.73}  Solve the initial value problem and graph the solution.

$y'''+2y''+y'+2y=30\cos x-10\sin x,
\quad y(0)=3,\quad y'(0)=-4,\quad y''(0)=16$
\end{problem}

\begin{problem}\label{exer:9.3.74}  Solve the initial value problem and graph the solution.

$y^{(4)}-3y'''+5y''-2y'=-2e^x(\cos x-\sin
x),\; y(0)=2,\; y'(0)=0,\; y''(0)~=~-1, \; y'''(0)=-5$
\end{problem}

\begin{problem}\label{exer:9.3.75}
Prove:
A function $y$ is a solution of the constant coefficient
nonhomogeneous equation
$$
a_0y^{(n)}+a_1y^{(n-1)}+\cdots+a_ny=e^{\alpha x}G(x)
\text{(A)}
$$
if and only if $y=ue^{\alpha x}$, where $u$ satisfies the differential
equation
$$
a_0u^{(n)}+\frac{p^{(n-1)}(\alpha)}{(n-1)!}u^{(n-1)}+
\frac{p^{(n-2)}(\alpha)}{(n-2)!}u^{(n-2)}+\cdots+p(\alpha)u=G(x)
\text{(B)}
$$
and
$$
p(r)=a_0r^n+a_1r^{n-1} + \cdots + a_n
$$
is the characteristic polynomial of the complementary equation
$$
a_0y^{(n)}+a_1y^{(n-1)}+\cdots+a_ny=0.
$$
\end{problem}

\begin{problem}\label{exer:9.3.76}
Prove:
\begin{enumerate}
\item % (a)
The equation
$$
\begin{array}{lcl}
a_0u^{(n)}&+&\frac{p^{(n-1)}(\alpha)}{(n-1)!}u^{(n-1)}+
\frac{p^{(n-2)}(\alpha)}{(n-2)!}u^{(n-2)}+\cdots+p(\alpha)u\\
&=&\left(p_0+p_1x+\cdots+p_kx^k\right)\cos
\omega x\\&&\,+ \left(q_0+q_1x+\cdots+q_kx^k\right)\sin\omega x
\end{array}
\text{(A)}
$$
has a particular solution of the form
$$
u_p=x^m\left(u_0+u_1x+\cdots+u_kx^k\right)\cos\omega x+
\left(v_0+v_1x+\cdots+v_kx^k\right)\sin\omega x.
$$
\item % (b)
If $\lambda+i\omega$ is a zero of $p$ with multiplicity $m\ge1$, then
(A) can be written as
$$
a(u''+\omega^2 u)=
\left(p_0+p_1x+\cdots+p_kx^k\right)\cos\omega x+
\left(q_0+q_1x+\cdots+q_kx^k\right)\sin\omega x,
$$
which has a particular solution of the form
$$
u_p=U(x)\cos\omega x+V(x)\sin\omega x,
$$
where
$$
U(x)=u_0x+u_1x^2+\cdots+u_kx^{k+1},\,V(x)=v_0x+v_1x^2+\cdots+v_kx^{k+1}
$$
and
$$
\begin{array}{rcl}
a(U''(x)+2\omega V'(x))&=&p_0+p_1x+\cdots+p_kx^k\\
a(V''(x)-2\omega U'(x))&=&q_0+q_1x+\cdots+q_kx^k.
\end{array}
$$
\end{enumerate}
\end{problem}


\end{document}