\documentclass{ximera}
\input{../preamble.tex}

\title{Exercises} \license{CC BY-NC-SA 4.0}

\begin{document}

\begin{abstract}
\end{abstract}
\maketitle

\begin{onlineOnly}
\section*{Exercises}
\end{onlineOnly}

\begin{problem}\label{exer:2.4.1} Solve the given Bernoulli
equation.
$$y'+y=y^2$$
\end{problem}

\begin{problem}\label{exer:2.4.2} Solve the given Bernoulli
equation.
$$7xy'-2y=-\frac{x^2}{y^6}$$



\begin{solution}
     Rewrite as $y'-\frac{2}{7x}y=-\frac{x}{7y^6}$. Then
$\frac{y_1'}{ y_1}=\frac{2}{7x}$;\,
$\ln|y_1|=\frac{2}{7}\ln|x|=\ln|x|^{2/7}$;\,
$y_1=x^{2/7}$;\,
$y=ux^{2/7}$;\,
$u'x^{2/7}=-\frac{1}{7u^6x^{5/7}}$;\,
$u^6u'=-\frac{1}{7x}$;\,
$\frac{u^7}{7}=-\frac{1}{7}\ln|x|+\frac{c}{7}$;\,
$u=(c-\ln|x|)^{1/7}$;\,
$y=x^{2/7}(c-\ln|x|)^{1/7}$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.3} Solve the given Bernoulli
equation.
$$x^2y'+2y=2e^{1/x}y^{1/2}$$
\end{problem}

\begin{problem}\label{exer:2.4.4} Solve the given Bernoulli
equation.
$$(1+x^2)y'+2xy=\frac{1}{(1+x^2)y}$$



\begin{solution}
    Rewrite as $y'+\frac{2x}{1+x^2}y=\frac{1}{(1+x^2)^2y}$.
Then
$\frac{y_1'}{ y_1}=-\frac{2x}{1+x^2}$;\,
$\ln|y_1|=-\ln(1+x^2)$;\,
$y_1=\frac{1}{1+x^2}$;\,
$y=\frac{u}{1+x^2}$;\,
$\frac{u'}{1+x^2}=\frac{1}{ u(1+x^2)}$;\,
$u'u=1$;\,
$\frac{u^2}{2}=x+\frac{c}{2}$;\,
$u=\pm\sqrt{2x+c}$;\,
$y=\pm\frac{\sqrt{2x+c}}{1+x^2}$.
\end{solution}
\end{problem}

 \begin{problem}\label{exer:2.4.5} Find all solutions. Also, plot a direction field and some integral curves on the indicated rectangular region.
$$y'-xy=x^3y^3;    \quad \{-3\leq x\leq 3, -2\leq y\leq 2\}$$
\end{problem}

\begin{problem}\label{exer:2.4.6} Find all solutions. Also, plot a direction field and some integral curves on the indicated rectangular region.
$$y'-\frac{1+x}{3x}y=y^4;   \quad \{-2\leq x\leq 2,-2\leq y \leq 2\}$$



\begin{solution}
    $\frac{y_1'}{ y_1}=\frac{1}{3}\left(\frac{1}{ x}+1\right)$;\,
$\ln|y_1|=\frac{1}{3}(\ln|x|+x)$;\,
$y_1=x^{1/3}e^{x/3}$;\,
$y=ux^{1/3}e^{x/3}$;\,
$u'x^{1/3}e^{x/3}=x^{4/3}e^{4x/3}u^4$;\,
$\frac{u'}{ u^4}=xe^x$;\,
$-\frac{1}{3u^3}=(x-1)e^x-\frac{c}{3}$;\,
$u=\frac{1}{[3(1-x)e^x+c]^{1/3}}$;\,
$y=\left[\frac{x}{3(1-x)+ce^{-x}}\right]^{1/3}$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.7} Solve the initial value problem.
$$y'-2y=xy^3,\quad y(0)=2\sqrt2$$
\end{problem}

\begin{problem}\label{exer:2.4.8} Solve the initial value problem.
$$y'-xy=xy^{3/2},\quad y(1)=4$$



\begin{solution}
    $\frac{y_1'}{ y_1}=x$;\,
$\ln|y_1|=\frac{x^2}{2}$;\,
$y_1=e^{x^2/2}$;\,
$y=ue^{x^2/2}$;\,
$u'e^{x^2/2}=xu^{3/2}e^{3x^2/4}$;\,
$\frac{u'}{ u^{3/2}}=xe^{x^2/4}$;\,
(A) $-\frac{2}{ u^{1/2}}=2e^{x^2/4}+2c$;\,
$u^{1/2}=-\frac{1}{ c+e^{x^2/4}}$;\,
$u=\frac{1}{(c+e^{x^2/4})^2}$;\,
$y=\frac{1}{(1+ce^{-x^2/4})^2}$. Because of (A) we must choose $c$
so that $y(1)=4$ and $1+ce^{-1/4}<0$. This implies that
$c=-3e^{1/4}$;\,
$y=\left[1-\frac{3}{2}e^{-(x^2-1)/4}\right]^{-2}$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.9} Solve the initial value problem.
$$xy'+y=x^4y^4,\quad y(1)=1/2$$
\end{problem}

\begin{problem}\label{exer:2.4.10} Solve the initial value problem.
$$y'-2y=2y^{1/2},\quad y(0)=1$$



\begin{solution}
    $\frac{y_1'}{ y_1}=2$;\,
$\ln|y_1|=2x$;\,
$y_1=e^{2x}$;\,
$y=ue^{2x}$;\,
$u'e^{2x}=2u^{1/2}e^x$;\,
$u^{-1/2}u'=2e^{-x}$;\,
$2u^{1/2}=-2e^{-x}+2c$;\,
$u^{1/2}=c-e^{-x}>0$;\,
$y(0)=1\Rightarrow u(0)=1\Rightarrow
c=2$;\,
$u=(2-e^{-x})^2$;\,
$y=(2e^x-1)^2$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.11} Solve the initial value problem. 
$$y'-4y=\frac{48x}{y^2},\quad y(0)=1$$
\end{problem}

\begin{problem}\label{exer:2.4.12} Solve the initial value problem and graph the solution. 
$$x^2y'+2xy=y^3,\quad y(1)=1/\sqrt2$$



\begin{solution}
    Rewrite as $y'+\frac{2}{ x}y=\frac{y^3}{ x^2}$. Then
$\frac{y_1'}{ y_1}=-\frac{2}{ x}$;\,
$\ln|y_1|=-2\ln|x|=\ln x^{-2}$;\,
$y_1=\frac{1}{ x^2}$;\,
$y=\frac{u}{ x^2}$;\,
$\frac{u'}{ x^2}=\frac{u^3}{ x^8}$;\,
$\frac{u'}{ u^3}=\frac{1}{ x^6}$;\,
$-\frac{1}{2u^2}=-\frac{1}{5x^5}+c$;\,
$y(1)=\frac{1}{\sqrt2}\Rightarrow u(1)=\frac{1}{\sqrt2}\Rightarrow
c=-\frac{4}{5}$;\,
$u=\left[\frac{5x^5}{2(1+4x^5)}\right]^{1/2}$;\,
$y=\left[\frac{5x}{2(1+4x^5)}\right]^{1/2}$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.13} Solve the initial value problem and graph the solution.
$$y'-y=xy^{1/2},\quad y(0)=4$$
\end{problem}

\begin{problem}\label{exer:2.4.14}
You may have noticed that the logistic equation
$$
P'=aP(1-\alpha P)
$$
from Verhulst's model for population growth can be written in
Bernoulli form as
$$
P'-aP=-a\alpha P^2.
$$
This isn't particularly interesting, since the logistic equation is
separable. So let's consider a more complicated model,
where
$a$ is a positive constant and $\alpha$ is a positive continuous
function of $t$ on $[0,\infty)$. The equation for this model is
$$
P'-aP=-a\alpha(t) P^2,
$$
a non-separable Bernoulli equation.
\begin{enumerate}
\item % (a)
Assuming that $P(0)=P_0>0$, find $P$ for $t>0$. 
\begin{hint}
Express your result in terms of the integral
 $\int_0^t\alpha(\tau)e^{a\tau}\,d\tau$.
 \end{hint}
\item % (b)
Verify that your result reduces to the known results
for the Malthusian model where  $\alpha=0$, and the Verhulst model
where $\alpha$ is a nonzero constant.
\item % (c)
Assuming that
$$
\lim_{t\to\infty}e^{-at}\int_0^t\alpha(\tau)e^{a\tau}\,d\tau=L
$$
exists (finite or infinite), find $\lim_{t\to\infty}P(t)$.
\end{enumerate}



\begin{solution}
    $P=ue^{at}$;\ $u'e^{at}=-a\alpha u^2e^{2at}$;\ $\frac{u'}{
u^2}=-a\alpha e^{at}$;\ $-\frac{1}{
u}=-a\int_0^t\alpha(\tau)e^{a\tau}\,d\tau-\frac{1}{ P_0}$;\;
$P=\frac{P_0e^{at}}{1+aP_0\int_0^t\alpha(\tau)e^{a\tau}\,d\tau}$,
which can also be written as
$P=\frac{P_0}{
e^{-at}+aP_0e^{-at}\int_0^t\alpha(\tau)e^{a\tau}\,d\tau}$.
Therefore, $\lim_{t\to\infty}P(t)=
\left\{\begin{array}{rl}\infty&\mbox{ if }L=0,\\0&\mbox{ if
}L=\infty,\\ 1/aL&\mbox{ if } 0<L<\infty.\end{array}\right.$
\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.55}
A generalized
\href{http://http://www-history.mcs.st-and.ac.uk/Indexes/Riccati.html}{Riccati equation} is of the form
\begin{equation}\label{eqA:2.4.55}
y'=P(x)+Q(x)y+R(x)y^2.
\end{equation}

(If $R\equiv-1$,  (\ref{eqA:2.4.55}) is a
\href{http://http://www-history.mcs.st-and.ac.uk/Indexes/Riccati.html}
{Riccati equation}.) Let $y_1$ be a known solution and $y$ an arbitrary solution of (\ref{eqA:2.4.55}). Let $z=y-y_1$. Show that $z$ is a
solution of a Bernoulli equation with $n=2$.
\end{problem}

\begin{problem}\label{exer:2.4.56} Given that $y_1$ is a solution of the given equation, use the method suggested by Exercise \ref{exer:2.4.55} to find other solutions.
$$y'=1+x - (1+2x)y+xy^2;  \quad    y_1=1$$



\begin{solution}
    If $y=z+1$, then
$z'+z=xz^2$;\;
$z=ue^{-x}$;\;
$u'e^{-x}=xu^2e^{-2x}$;\;
$\frac{u'}{ u^2}=xe^{-x}$;\;
$-\frac{1}{ u}=-e^{-x}(x+1)-c$;\;
$u=\frac{1}{ e^{-x}(x+1)+c}$;\;
$z=\frac{1}{ x+1+ce^x}$;\;
$y=1+\frac{1}{ x+1+ce^x}$.

\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.57} Given that $y_1$ is a solution of the given equation, use the
method suggested by Exercise \ref{exer:2.4.55} to find other solutions.
$$y'=e^{2x}+(1-2e^x)y+y^2;  \quad    y_1=e^x$$
\end{problem}

\begin{problem}\label{exer:2.4.58} Given that $y_1$ is a solution of the given equation, use the
method suggested by Exercise \ref{exer:2.4.55} to find other solutions.
$$xy'=2-x+(2x-2)y-xy^2;  \quad    y_1=1$$



\begin{solution}
    If $y=z+1$, then
$z'+\frac{2}{ x}z=z^2$;\;
$z_1=\frac{1}{ x^2}$;\;
$z=\frac{u}{ x^2}$;\;
$\frac{u'}{ x^2}=\frac{u^2}{ x^4}$;\;
$\frac{u'}{ u^2}=\frac{1}{ x^2}$;\;
$-\frac{1}{ u}=-\frac{1}{ x}+c=-\frac{1-cx}{ x}$;\;
$u=-\frac{x}{ 1-cx}$;\;
$z=-\frac{1}{ x(1-cx)}$;\;
$y=1-\frac{1}{ x(1-cx)}$.

\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.59} Given that $y_1$ is a solution of the given equation, use the
method suggested by Exercise \ref{exer:2.4.55} to find other solutions.
$$xy'=x^3+(1-2x^2)y+xy^2;  \quad    y_1=x$$
\end{problem}

\end{document}