\documentclass{ximera}
\input{../preamble.tex}

\title{Exercises} \license{CC BY-NC-SA 4.0}

\begin{document}

\begin{abstract}
\end{abstract}
\maketitle

\begin{onlineOnly}
\section*{Exercises}
\end{onlineOnly}


 \begin{problem}\label{exer:10.7.1}
 Find a particular solution.
 
 $ {\bf y}'=\begin{bmatrix}-1&-4\\-1&-1\end{bmatrix}{\bf
y}+\begin{bmatrix}21e^{4t}\\8e^{-3t}\end{bmatrix}$
\end{problem}

 \begin{problem}\label{exer:10.7.2}
 Find a particular solution.
 
 $ {\bf
y}'=\frac{1}{5}\begin{bmatrix}-4&3\\-2&-11\end{bmatrix}{\bf
y}+\begin{bmatrix}50e^{3t}\\10e^{-3t}\end{bmatrix}$

\begin{solution}
    $Y=\begin{bmatrix}
-3e^{-t}&-e^{-2t}\\e^{-t}&2e^{-2t}\end{bmatrix}$;
${\bf u}'=Y^{-1}{\bf f}=
\frac{1}{5}\begin{bmatrix}-2e^t&-e^t
\\e^{2t}&3e^{2t}\end{bmatrix}
\begin{bmatrix}
50e^{37}\\10e^{-3t}
\end{bmatrix}=
\begin{bmatrix}-20e^{4t}-2e^{-2t}\\
10e^{5t}+6e^{-t}\end{bmatrix}$;
${\bf u}=
\begin{bmatrix}e^{-2t}-5e^{4t}\\
2e^{5t}-6e^{-t}\end{bmatrix}$;
${\bf y}_p=Y{\bf u}=
\begin{bmatrix}13e^{3t}+3e^{-3t}\\
-e^{3t}-11e^{-3t}\end{bmatrix}$.

\end{solution}
\end{problem} 


 \begin{problem}\label{exer:10.7.3}
 Find a particular solution.
 
 $ {\bf y}'=\begin{bmatrix}1&2\\2&1\end{bmatrix}{\bf
y}+\begin{bmatrix}1\\t\end{bmatrix}$
\end{problem}

 \begin{problem}\label{exer:10.7.4}
 Find a particular solution.
 
 $ {\bf y}'=\begin{bmatrix}-4&-3\\6&5\end{bmatrix}{\bf
y}+\begin{bmatrix}2\\-2e^t\end{bmatrix}$

\begin{solution}
    $Y=\begin{bmatrix}
-e^{2t}&-e^{-t}\\2e^{2t}&e^{-t}\end{bmatrix}$;
${\bf u}'=Y^{-1}{\bf f}=
\begin{bmatrix}e^{-2t}&e^{-2t}\\
-2e^t&-e^t\end{bmatrix}
\begin{bmatrix}2\\-2e^t\end{bmatrix}=
\begin{bmatrix}2e^{-2t}-2e^{-t}\\2e^{2t}-4e^t\end{bmatrix}$;
${\bf u}=
\begin{bmatrix}2e^{-t}-e^{-2t}\\
e^{2t}-4e^t\end{bmatrix}$;
${\bf y}_p=Y{\bf u}=
\begin{bmatrix}5-3e^t\\5e^t-6\end{bmatrix}$.
\end{solution}
 \end{problem}


 \begin{problem}\label{exer:10.7.5}
 Find a particular solution.
 
 $ {\bf y}'=\begin{bmatrix}-6&-3\\1&-2\end{bmatrix}{\bf
y}+\begin{bmatrix}4e^{-3t}\\4e^{-5t}\end{bmatrix}$
\end{problem}

 \begin{problem}\label{exer:10.7.6}
 Find a particular solution.
 
 $ {\bf y}'=\begin{bmatrix}0&1\\-1&0\end{bmatrix}{\bf
y}+\begin{bmatrix}1\\t\end{bmatrix}$

\begin{solution}
    $Y=\begin{bmatrix}\sin t&-\cos t\\
\cos t&\sin t\end{bmatrix}$;
${\bf u}'=Y^{-1}{\bf f}=
\begin{bmatrix}\sin t&\cos t\\-\cos t&\sin t\end{bmatrix}
\begin{bmatrix}1\\t\end{bmatrix}=
\begin{bmatrix}t\cos t+\sin t\\t\sin t-\cos
t\end{bmatrix}$;
${\bf u}=
\begin{bmatrix}t\sin t\\-t\cos t\end{bmatrix}$;
${\bf y}_p=Y{\bf u}=
\begin{bmatrix}t\\0\end{bmatrix}$.
\end{solution}
 \end{problem}


 \begin{problem}\label{exer:10.7.7}
 Find a particular solution.
 
 $ {\bf y}'=\begin{bmatrix}3&1&-1\\3&5&1\\-6&2&4\end{bmatrix}{\bf
y}+\begin{bmatrix}3\\6\\3\end{bmatrix}$
\end{problem}

 \begin{problem}\label{exer:10.7.8}
 Find a particular solution.
 
 $ {\bf
y}'=\begin{bmatrix}3&-1&-1\\-2&3&2\\4&-1&-2\end{bmatrix}{\bf
y}+\begin{bmatrix}1\\e^t\\e^t\end{bmatrix}$

\begin{solution}
    $Y=\begin{bmatrix}
e^{3t}&e^{2t}&e^{-t}\\-e^{3t}&0&-3e^{-t}\\e^{3t}&e^{2t}&7
e^{-t}\end{bmatrix}$;
${\bf u}'=Y^{-1}{\bf f}=
\frac{1}{6}\begin{bmatrix}
3e^{-3t}&-6e^{-3t}&-3e^{-3t}\\4e^{-2t}&6e^{-2t}&2e^{-2t}\\
-e^t&0&e^t\end{bmatrix}
\begin{bmatrix}1\\e^t\\e^t\end{bmatrix}=
\frac{1}{6}\begin{bmatrix}
3e^{-3t}-9e^{-2t}\\8e^{-t}+4e^{-2t}\\e^{2t}-e^t\end{bmatrix}$;
${\bf u}=
\frac{1}{12}\begin{bmatrix}9e^{-2t}-2e^{-3t}\\
-16e^{-t}-4e^{-2t}\\e^{2t}-2e^t\end{bmatrix}$;
${\bf y}_p=Y{\bf u}=
-\frac{1}{6}\begin{bmatrix}3e^t+4\\6e^t-4\\10
\end{bmatrix}$.
\end{solution}
\end{problem}
 

 \begin{problem}\label{exer:10.7.9}
 Find a particular solution.

$ {\bf y}'=\begin{bmatrix}-3&2&2\\2&-3&2\\2&2&-3\end{bmatrix}{\bf
y}+\begin{bmatrix}e^t\\e^{-5t}\\e^t\end{bmatrix}$
\end{problem}

 \begin{problem}\label{exer:10.7.10}
 Find a particular solution.
 
 $ {\bf
y}'=\frac{1}{3}\begin{bmatrix}1&1&-3\\-4&-4&3\\-2&1&0\end{bmatrix}{\bf
y}+\begin{bmatrix}e^t\\e^t\\e^t\end{bmatrix}$

\begin{solution}
    $Y=\begin{bmatrix}
-e^t&e^{-t}&te^{-t}\\e^t&-e^{-t}&3e^{-t}-te^{-t}\\
e^t&e^{-t}&te^{-t}\end{bmatrix}$;
${\bf u}'=Y^{-1}{\bf f}=
\frac{1}{18}\begin{bmatrix}
-9e^{-t}&0&9e^{-t}\\3e^t(3-2t)&-6te^t&9e^t\\
6e^t&6e^t&0\end{bmatrix}
\begin{bmatrix}e^t\\e^t\\e^t\end{bmatrix}=
\frac{1}{3}\begin{bmatrix}0\\
e^{2t}(3-2t)\\2e^{2t}\end{bmatrix}$;
${\bf u}=
\frac{1}{3}\begin{bmatrix}0\\
e^{2t}(2-t)\\e^{2t}\end{bmatrix}$;
${\bf y}_p=Y{\bf u}=
\frac{1}{3}\begin{bmatrix}
2e^t\\e^t\\2e^t\end{bmatrix}$.
\end{solution}
\end{problem}

 \begin{problem}\label{exer:10.7.11}
 Find a particular solution, given that $Y$ is a fundamental matrix for the complementary system.

$ {\bf y}'=\frac{1}{t}\begin{bmatrix}1&t\\-t&1\end{bmatrix}{\bf y}
+t\begin{bmatrix}
\cos t\\\sin t\end{bmatrix};  \quad Y=t\begin{bmatrix}\cos
t&\sin t\\-\sin t&\cos t
\end{bmatrix}$
\end{problem}

 \begin{problem}\label{exer:10.7.12}
 Find a particular solution, given that $Y$ is a fundamental matrix for the complementary system.
 
$ {\bf y}'=\frac{1}{t}\begin{bmatrix}1&t\\t&1\end{bmatrix}{\bf y}
+\begin{bmatrix}
t\\t^2\end{bmatrix};  \quad Y=t\begin{bmatrix}
e^t&e^{-t}\\e^t&-e^{-t}
\end{bmatrix}$

\begin{solution}
    ${\bf u}'=Y^{-1}{\bf f}=
\frac{1}{2t}\begin{bmatrix}
e^{-t}&e^{-t}\\e^t&-e^t\end{bmatrix}
\begin{bmatrix}t\\t^2\end{bmatrix}=
\frac{1}{2}\begin{bmatrix}
e^{-t}(t+1)\\e^t(1-t)\end{bmatrix}$;
${\bf u}=
\frac{1}{2}\begin{bmatrix}
-e^{-t}(t+2)\\e^t(2-t)\end{bmatrix}$;
${\bf y}_p=Y{\bf u}=
\frac{t}{2}\begin{bmatrix}
e^t&e^{-t}\\e^t&-e^{-t}\end{bmatrix}
\begin{bmatrix}-e^{-t}(t+2)\\
e^t(2-t)\end{bmatrix}=
-\begin{bmatrix}t^2\\2t\end{bmatrix}$.
\end{solution}
\end{problem}

 \begin{problem}\label{exer:10.7.13}
 Find a particular solution, given that $Y$ is a fundamental matrix for the complementary system.
 
$ {\bf y}'=\frac{1}{t^2-1}\begin{bmatrix}t&-1\\-1&t\end{bmatrix}{\bf y}+
t\begin{bmatrix}
1\\-1\end{bmatrix};  \quad Y=\begin{bmatrix}
t&1\\1&t
\end{bmatrix}$
\end{problem}

 \begin{problem}\label{exer:10.7.14}
 Find a particular solution, given that $Y$ is a fundamental matrix for the complementary system.
 
$ {\bf y}'=\frac{1}{3}\begin{bmatrix}1&-2e^{-t}\\2e^t&-1\end{bmatrix}{\bf y}+
\begin{bmatrix}e^{2t}
\\e^{-2t}\end{bmatrix};  \quad Y=\begin{bmatrix}
2&e^{-t}\\e^t&2
\end{bmatrix}$

\begin{solution}
    ${\bf u}'=Y^{-1}{\bf f}=
\frac{1}{3}\begin{bmatrix}2&-e^{-t}\\-e^t&2\end{bmatrix}
\begin{bmatrix}e^{2t}\\e^{-2t}\end{bmatrix}=
\frac{1}{3}\begin{bmatrix}
2e^{2t}-e^{-3t}\\2e^{-2t}-e^{3t}\end{bmatrix}$;
${\bf u}=
\frac{1}{9}\begin{bmatrix}
3e^{2t}+e^{-3t}\\-e^{3t}-3e^{-2t}\end{bmatrix}$;
${\bf y}_p=Y{\bf u}=
\frac{1}{9}\begin{bmatrix}
2&e^{-t}\\e^t&2\end{bmatrix}
\begin{bmatrix}
3e^{2t}+e^{-3t}\\-e^{3t}-3e^{-2t}\end{bmatrix}=
{1}{9}\begin{bmatrix}
5e^{2t}-e^{-3t}\\e^{3t}-5e^{-2t}\end{bmatrix}$.
\end{solution}
\end{problem}

 \begin{problem}\label{exer:10.7.15}
 Find a particular solution, given that $Y$ is a fundamental matrix for the complementary system.
 
$ {\bf y}'=\frac{1}{2t^4}\begin{bmatrix}3t^3&t^6\\1&-3t^3\end{bmatrix}{\bf y}+
\frac{1}{t}\begin{bmatrix}
t^2\\1\end{bmatrix};  \quad Y=\frac{1}{t^2}\begin{bmatrix}
t^3&t^4\\-1&t
\end{bmatrix}$
\end{problem}

 \begin{problem}\label{exer:10.7.16}
 Find a particular solution, given that $Y$ is a fundamental matrix for the complementary system.
 
$ {\bf y}'=
\begin{bmatrix} \frac{1}{t-1}
&- \frac{e^{-t}}{t-1}\\ \frac{e^t}{t+1}&\frac{1}{t+1}\end{bmatrix}{\bf y}+\begin{bmatrix}
t^2-1\\t^2-1\end{bmatrix};  \quad Y=\begin{bmatrix}
t&e^{-t}\\e^t&t
\end{bmatrix}$

\begin{solution}
    ${\bf u}'=Y^{-1}{\bf f}=
\frac{1}{ t^2-1}\begin{bmatrix}
t&-e^{-t}\\-e^t&t\end{bmatrix}
\begin{bmatrix}t^2-1\\t^2-1\end{bmatrix}=
\begin{bmatrix}t-e^{-t}\\t-e^t\end{bmatrix}$;
${\bf u}=
\frac{1}{2}\begin{bmatrix}
2e^{-t}+t^2\\t^2-2e^t\end{bmatrix}$;
${\bf y}_p=Y{\bf u}
=\frac{1}{2}\begin{bmatrix}t&e^{-t}\\e^t&t\end{bmatrix}
\begin{bmatrix}
2e^{-t}+t^2\\t^2-2e^t\end{bmatrix}=
\frac{1}{2}\begin{bmatrix}
te^{-t}(t+2)+t^3-2\\te^t(t-2)+t^3+2\end{bmatrix}$.
\end{solution}
\end{problem}

 \begin{problem}\label{exer:10.7.17}
 Find a particular solution, given that $Y$ is a fundamental matrix for the complementary system.
 
$ {\bf y}'=\frac{1}{t}\begin{bmatrix}1&1&0\\0&2&1\\-2&2&2\end{bmatrix}{\bf y}
+\begin{bmatrix}1\\2\\1\end{bmatrix}
\quad Y=\begin{bmatrix}
t^2&t^3&1\\t^2&2t^3&-1\\0&2t^3&2
\end{bmatrix}$
\end{problem}

 \begin{problem}\label{exer:10.7.18}
 Find a particular solution, given that $Y$ is a fundamental matrix for the complementary system.
 
$ {\bf y}'=
\begin{bmatrix}3&e^t&e^{2t}\\e^{-t}&2&e^t\\e^{-2t}&e^{-t}&1\end{bmatrix}{\bf
y}
+\begin{bmatrix}
e^{3t}\\0 \\ 0\end{bmatrix};  \quad Y=\begin{bmatrix}
e^{5t}&e^{2t}&0\\e^{4t}&0&e^t\\e^{3t}&-1&-1
\end{bmatrix}$

\begin{solution}
    ${\bf u}'=Y^{-1}{\bf f}=
\frac{1}{3}\begin{bmatrix}e^{-5t}&e^{-4t}&e^{-3t}\\
2e^{-2t}&-e^{-t}&-1\\-e^{-2t}&2e^{-t}&-1\end{bmatrix}
\begin{bmatrix}e^t\\0\\0\end{bmatrix}=
\frac{1}{3}\begin{bmatrix}
e^{-4t}\\2e^{-t}\\-e^{-t}\end{bmatrix}$;
${\bf u}=
\frac{1}{12}\begin{bmatrix}
-e^{-4t}\\-8e^{-t}\\4e^{-t}\end{bmatrix}$;
${\bf y}_p=Y{\bf u}=
\frac{1}{12}\begin{bmatrix}e^{5t}&e^{2t}&0\\
e^{4t}&0&e^t\\e^{3t}&-1&-1\end{bmatrix}
\begin{bmatrix}
-e^{-4t}\\-8e^{-t}\\4e^{-t}\end{bmatrix}=
\frac{1}{4}\begin{bmatrix}-3e^t\\
1\\e^{-t}\end{bmatrix}$.
\end{solution}
\end{problem}

 \begin{problem}\label{exer:10.7.19}
 Find a particular solution, given that $Y$ is a fundamental matrix for the complementary system.
 
$ {\bf y}'=\frac{1}{t}\begin{bmatrix}1&t&0\\0&1&t\\0&-t&1\end{bmatrix}{\bf y}
+\begin{bmatrix}
t\\t \\t \end{bmatrix};  \quad Y=t\begin{bmatrix}
1&\cos t&\sin t \\0&-\sin t&\cos t\\0&-\cos t&-\sin t
\end{bmatrix}$
\end{problem}

 \begin{problem}\label{exer:10.7.20}
 Find a particular solution, given that $Y$ is a fundamental matrix for the complementary system.
 
$ {\bf y}'=-\frac{1}{t}\begin{bmatrix}e^{-t}&-t&1-e^{-t}\\e^{-t}&1&-t-e^{-t}\\e^{-t}&-t
&1-e^{-t}\end{bmatrix}{\bf
y}
+\frac{1}{t}\begin{bmatrix}
e^t\\0 \\e^t \end{bmatrix};  \quad Y=\frac{1}{t}\begin{bmatrix} e^t&e^{-t}&t\\e^t&-e^{-t}&e^{-t}\\e^t&e^{-t}&0
\end{bmatrix}$

\begin{solution}
    ${\bf u}'=Y^{-1}{\bf f}=
\frac{1}{2t}\begin{bmatrix}
-e^{-2t}&te^{-t}&te^{-t}+e^{-2t}\\1&-te^t&te^t-1\\
2&0&-2\end{bmatrix}
\begin{bmatrix}e^t\\0\\e^t\end{bmatrix}=
\begin{bmatrix}1\\e^{2t}\\0\end{bmatrix}$;
${\bf u}=
\frac{1}{4}\begin{bmatrix}
2t\\e^{2t}\\0\end{bmatrix}$;
${\bf y}_p=Y{\bf u}=
Y=\frac{1}{4t}\begin{bmatrix}e^t&e^{-t}&t\\
e^t&-e^{-t}&e^{-t}\\e^t&e^{-t}&0\end{bmatrix}
\begin{bmatrix}2t\\e^{2t}\\0\end{bmatrix}=
{e^{t}}{4t}\begin{bmatrix}
2t+1\\2t-1\\2t+1\end{bmatrix}$.
\end{solution}
\end{problem}

 \begin{problem}\label{exer:10.7.21}  Prove Theorem~\ref{thmtype:10.7.1}.
 \end{problem}

 \begin{problem}\label{exer:10.7.22}
\begin{enumerate}
\item % (a)
Convert the scalar equation
$$
P_0(t)y^{(n)}+P_1(t)y^{(n-1)}+\cdots+P_n(t)y=F(t)
\quad\quad\text{(A)}
$$
into an equivalent $n\times n$ system
$$
{\bf y}'=A(t){\bf y}+{\bf f}(t).
\quad\quad\text{(B)}
$$
\item % (b)
Suppose  (A) is normal on an interval $(a,b)$ and
$\{y_1,y_2,\dots,y_n\}$ is a fundamental set of solutions of
$$
P_0(t)y^{(n)}+P_1(t)y^{(n-1)}+\cdots+P_n(t)y=0
\quad\quad\text{(C)}
$$
on $(a,b)$. Find a corresponding fundamental matrix $Y$ for
$$
{\bf y}'=A(t){\bf y}
\quad\quad\text{(D)}
$$
on $(a,b)$ such that
$$
y=c_1y_1+c_2y_2+\cdots+c_ny_n
$$
is a solution of  (C) if and only if
${\bf y}=Y{\bf c}$ with
$$
{\bf c}=\begin{bmatrix}c_1\\c_2\\\vdots\\c_n\end{bmatrix}
$$
is a solution of  (D).
\item % (c)
Let $y_p=u_1y_1+u_1y_2+\cdots+u_ny_n$ be a particular solution of
 (A), obtained by the method of variation of parameters for scalar
equations as given in Section~9.4, and define
$$
{\bf u}=\begin{bmatrix}u_1\\u_2\\\vdots\\u_n\end{bmatrix}.
$$
Show that ${\bf y}_p=Y{\bf u}$ is a  solution of  (B).

\begin{solution}
    If ${\bf y}_p=Y{\bf u}$, then ${\bf y}_p'=Y'{\bf u}+Y{\bf u}'=AY{\bf
u}+Y{\bf u}'$, so (E) ${\bf y}_p'=A{\bf y_p}+Y{\bf u}'$. However, from
the derivation of the method of variation of parameters in
Section~9.4, $Y{\bf u}'={\bf f}$ as defined in the solution
of part (a). This and (E) imply the conclusion.
\end{solution}

\item % (d)
Let  ${\bf y}_p=Y{\bf u}$ be  a particular solution of  (B),
obtained by the method of variation of parameters for systems as given in
this section. Show that
 $y_p=u_1y_1+u_1y_2+\cdots+u_ny_n$ is a solution of
 (A).

 \begin{solution}
     Since $Y{\bf u}'={\bf f}$ with ${\bf f}$ as defined in the
solution of part(a), $u_1,u_2,\dots,u_n$ satisfy
the conditions required in the derivation of the method of
variation of parameters in Section~9.4; hence,
$y_p=c_1y_1+c_2y_2+\cdots+c_ny_n$ is a particular solution of (A).
 \end{solution}
\end{enumerate}
\end{problem}

 \begin{problem}\label{exer:10.7.23}
Suppose the $n\times n$ matrix function $A$ and the $n$--vector
function ${\bf f}$ are continuous on $(a,b)$. Let $t_0$ be in $(a,b)$,
let ${\bf k}$ be an arbitrary constant vector, and let $Y$ be a
fundamental matrix for the homogeneous system ${\bf y}'=A(t){\bf y}$.
Use variation of parameters to show that the solution of the initial
value problem
$$
{\bf y}'=A(t){\bf y}+{\bf f}(t),\quad {\bf y}(t_0)={\bf k}
$$
is
$$
{\bf y}(t)=Y(t)\left( Y^{-1}(t_0){\bf k}+\int_{t_0}^t
Y^{-1}(s){\bf f}(s)\, ds\right).
$$
\end{problem}

\end{document}