\documentclass{ximera}
\input{../preamble.tex}

\title{Exercises} \license{CC BY-NC-SA 4.0}

%NO ANSWERS ADDED YET
\begin{document}

\begin{abstract}
\end{abstract}
\maketitle

\begin{onlineOnly}
\section*{Exercises}
\end{onlineOnly}



\begin{problem}\label{exer:1.3.12} Construct a direction field and plot some integral curves in the indicated rectangular region.
$$y'=y(y-1);    \quad  \{-1\leq x\leq 2,\ -2\leq y\le2\}$$
\end{problem}

\begin{problem}\label{exer:1.3.13} Construct a direction field and plot some integral curves in the indicated rectangular region.
$$y'=2-3xy;    \quad  \{-1\leq x\leq 4,\ -4\leq y\leq 4\}$$
\end{problem}

\begin{problem}\label{exer:1.3.14} Construct a direction field and plot some integral curves in the indicated rectangular region.
$$y'=xy(y-1);    \quad  \{-2\leq x\leq 2,\ -4\leq y\leq 4\}$$
\end{problem}

\begin{problem}\label{exer:1.3.15} Construct a direction field and plot some integral curves in the indicated rectangular region.
$$y'=3x+y;    \quad  \{-2\leq x\leq 2,\ 0\leq y\leq 4\}$$
\end{problem}

\begin{problem}\label{exer:1.3.16} Construct a direction field and plot some integral curves in the indicated rectangular region.
$$y'=y-x^3;    \quad  \{-2\leq x\leq 2,\ -2\leq y\leq 2\}$$
\end{problem}

\begin{problem}\label{exer:1.3.17} Construct a direction field and plot some integral curves in the indicated rectangular region.
$$y'=1-x^2-y^2;    \quad  \{-2\leq x\leq 2,\ -2\leq y\leq 2\}$$
\end{problem}

\begin{problem}\label{exer:1.3.18} Construct a direction field and plot some integral curves in the indicated rectangular region.
$$y'=x(y^2-1);    \quad  \{-3\leq x\leq 3,\ -3\leq y\leq 2\}$$
\end{problem}

\begin{problem}\label{exer:1.3.19} Construct a direction field and plot some integral curves in the indicated rectangular region.
$$y'=\frac{x}{y(y^2-1)};    \quad  \{-2\leq x\leq 2,\ -2\leq y\leq 2\}$$
\end{problem}

\begin{problem}\label{exer:1.3.20} Construct a direction field and plot some integral curves in the indicated rectangular region.
$$y'=\frac{xy^2}{y-1};    \quad  \{-2\leq x\leq 2,\ -1\leq y\leq 4\}$$
\end{problem}

\begin{problem}\label{exer:1.3.21} Construct a direction field and plot some integral curves in the indicated rectangular region.
$$y'=\frac{x(y^2-1)}{y};    \quad  \{-1\leq x\leq 1,\ -2\leq y\leq 2\}$$
\end{problem}

\begin{problem}\label{exer:1.3.22} Construct a direction field and plot some integral curves in the indicated rectangular region.
$$y'=-\frac{x^2+y^2}{1-x^2-y^2};    \quad  \{-2\leq x\leq 2,\ -2\leq y\leq 2\}$$
\end{problem}

\begin{problem}\label{exer:1.3.23}
By suitably renaming the constants and  dependent variables
in the equations
\begin{equation}\label{eqA:1.3.23}
T' = -k(T-T_m)
\end{equation}
and
\begin{equation}\label{eqB:1.3.23}
G'=-\lambda G+r
\end{equation}
discussed in Section~1.2 in connection with Newton's
law of cooling and absorption of glucose in the body, we can
write both as
\begin{equation}\label{eqC:1.3.23}
y'=- ay+b,
\end{equation}
where $a$ is a positive constant and $b$ is an arbitrary
constant. Thus, (\ref{eqA:1.3.23}) is of the form (\ref{eqC:1.3.23}) with $y=T$, $a=k$, and
$b=kT_m$, and (\ref{eqB:1.3.23}) is of the form (\ref{eqC:1.3.23}) with $y=G$,
$a=\lambda$, and $b=r$. We'll encounter equations of the form
(\ref{eqC:1.3.23}) in many other applications in Chapter~2.

Choose a positive  $a$  and an
arbitrary  $b$.
 Construct a  direction field and plot some integral curves for (\ref{eqC:1.3.23}) in a rectangular region of the form
$$
\{0\leq t\leq T,\ c\leq y\leq d\}
$$
 of the $ty$-plane.
 Vary $T$, $c$, and $d$ until you discover a common property
of all the solutions of (\ref{eqC:1.3.23}). Repeat this experiment with various
choices of $a$ and $b$ until you can state
this property precisely in terms of $a$ and $b$.
\end{problem}

\begin{problem}\label{exer:1.3.24}
By suitably renaming the constants and  dependent variables
in the equations
\begin{equation}\label{eqA:1.3.24}
P'=aP(1-\alpha P)
\end{equation}
and
\begin{equation}\label{eqB:1.3.24}
I'=rI(S-I)
\end{equation}

discussed in Section~1.1 in connection with Verhulst's
population
model and the spread of an epidemic, we can write both in the form
\begin{equation}\label{eqC:1.3.24}
y'=ay-by^2,
\end{equation}
where $a$ and $b$ are positive constants. Thus, (\ref{eqA:1.3.24}) is of the form (\ref{eqC:1.3.24})
with $y=P$, $a=a$, and $b=a\alpha$, and (\ref{eqB:1.3.24}) is of the form (\ref{eqC:1.3.24}) with
$y=I$, $a=rS$, and $b=r$. In Chapter~2 we'll encounter
equations of the form
(\ref{eqC:1.3.24}) in
 other applications..

\begin{enumerate}
\item % (a)
Choose positive numbers $a$ and $b$. Construct a direction field and
plot some integral curves for (\ref{eqC:1.3.24}) in a rectangular region of the form
$$
\{0\leq t\leq T,\ 0\leq y\leq d\}
$$
of the $ty$-plane. Vary $T$ and $d$ until you discover a common
property of all solutions of (\ref{eqC:1.3.24}) with $y(0)>0$. Repeat this experiment
with various choices of $a$ and $b$ until you can state this property
precisely in terms of $a$ and $b$.

\item % (b)
Choose  positive numbers $a$ and $b$. Construct a direction field and
plot some integral curves for (\ref{eqC:1.3.24}) in a rectangular region of the form
$$
\{0\leq t\leq T,\ c\leq y\leq 0\}
$$
of the $ty$-plane. Vary $a$, $b$, $T$ and $c$ until you discover a
common property of all solutions of (\ref{eqC:1.3.24}) with $y(0)<0$.

You can verify your results later by doing Problem \ref{exer:2.2.27}.

\end{enumerate}
\end{problem}
\end{document}