\documentclass{ximera}
\input{../preamble.tex}

\title{Exercises} \license{CC BY-NC-SA 4.0}

\begin{document}

\begin{abstract}
\end{abstract}
\maketitle

\begin{onlineOnly}
\section*{Exercises}
\end{onlineOnly}

\begin{problem}\label{exer:6.2.1}
A 64 lb object stretches a spring 4 ft in equilibrium. It is attached
to a dashpot with damping constant $c=8$ lb-sec/ft. The object is
initially displaced 18 inches above equilibrium and given a downward
velocity of 4 ft/sec. Find its displacement and time--varying
amplitude for $t>0$.
\end{problem}

\begin{problem}\label{exer:6.2.2}  
A 16 lb weight is attached to a spring with natural length 5 ft. With
the weight attached, the spring measures 8.2 ft. The weight is
initially displaced 3 ft below equilibrium and given an upward
velocity of 2 ft/sec. Find and graph its displacement
 for $t>0$ if the medium resists the motion with a force of
one lb for each ft/sec of velocity. Also, find its time--varying amplitude.

\begin{solution}
    Since $k=\frac{mg}{\Delta l}=\frac{16}{3.2}=5$ the equation of
motion is $\frac{1}{2}y''+y'+5y=0$, or (A) $y''+2y'+10y=0$. The
characteristic polynomial of (A) is $p(r)=r^2+2r+10=(r+1)^2+9$.
Therefore,the general solution of (A) is
$y=e^{-t}(c_1\cos3t+c_2\sin3t)$, so
$y'=-y+3e^{-t}(-c_1\sin3t+c_2\cos3t)$. Now $y(0)=-3$ and
$y'(0)=2\Rightarrow c_1=-3$ and $2=3+3c_2$, or $c_2=-\frac{1}{3}$.
Therefore,$y=-e^{-t}\left(3\cos3t+{1}{3}\sin3t\right)$ ft.
The time--varying amplitude is $\frac{\sqrt{82}}{3}e^{-t}$ ft.
\end{solution}
\end{problem}

\begin{problem}\label{exer:6.2.3}  
An 8 lb weight stretches a spring 1.5 inches. It is attached to a
dashpot with damping constant $c$=8 lb-sec/ft. The weight is
initially displaced 3 inches above equilibrium and given an upward
velocity of 6 ft/sec. Find and graph its displacement for $t>0$.
\end{problem}

\begin{problem}\label{exer:6.2.4}
A 96 lb weight stretches a spring 3.2 ft in equilibrium. It is
attached to a dashpot with damping constant $c$=18 lb-sec/ft. The
weight is initially displaced 15 inches below equilibrium and given a
downward velocity of 12 ft/sec. Find its displacement for $t>0$.

\begin{solution}
    Since $k=\frac{mg}{\Delta l}=\frac{96}{3.2}=30$ the equation of
motion is $3y''+18y'+30y=0$, or (A) $y''+6y'+10y=0$. The
characteristic polynomial of (A) is $p(r)=r^2+6r+10=(r+3)^2+1$.
Therefore,the general solution of (A) is $y=e^{-3t}(c_1\cos t+c_2\sin
t)$, so $y'=-3y+e^{-3t}(-c_1\sin t+c_2\cos t)$. Now
$y(0)=-\frac{5}{4}$ and $y'(0)=-12\Rightarrow c_1=-\frac{5}{4}$
and $-12=\frac{15}{4}+c_2$, or $c_2=-\frac{63}{4}$. Therefore,
$y=-\frac{e^{-3t}}{4}(5\cos t+63\sin t)$ ft.
\end{solution}
\end{problem}

\begin{problem}\label{exer:6.2.5}
  A 16 lb weight  stretches a spring 6 inches in equilibrium.
It is attached to a damping mechanism with constant $c$.  Find all values
of $c$ such that the free vibration of the weight has infinitely many
oscillations.
\end{problem}

\begin{problem}\label{exer:6.2.6}
An 8 lb weight stretches a spring .32 ft. The weight is initially
displaced 6 inches above equilibrium and given an upward velocity of 4
ft/sec. Find its displacement for $t>0$ if the medium exerts a
damping force of 1.5 lb for each ft/sec of velocity.

\begin{solution}
    Since $k=\frac{mg}{\Delta l}=\frac{8}{.32}=25$ the equation of
motion is $\frac{1}{4}y''+\frac{3}{2}y'+25y=0$, or (A)
$y''+6y'+100y=0$. The characteristic polynomial of (A) is
$p(r)=r^2+6r+100=(r+3)^2+91$. Therefore,the general solution of (A) is
$y=e^{-3t}(c_1\cos\sqrt{91}t+c_2\sin\sqrt{91}t)$, so
$y'=-3y+\sqrt{91}e^{-3t}(-c_1\sin\sqrt{91}t+c_2\cos\sqrt{91}t)$. Now
$y(0)=\frac{1}{2}$ and $y'(0)=4\Rightarrow c_1=\frac{1}{2}$ and
$4=-\frac{3}{2}+\sqrt{91}c_2$, or $c_2=\frac{11}{2\sqrt{91}}$.
Therefore,$y=\frac{1}{2}e^{-3t}\left(\cos\sqrt{91}t+\frac{11}{
\sqrt{91}}\sin\sqrt{91}t\right)$ ft.
\end{solution}
\end{problem}

\begin{problem}\label{exer:6.2.7}
A 32 lb weight stretches a spring 2 ft in equilibrium. It is attached
to a dashpot with constant $c=8$ lb-sec/ft. The weight is initially
displaced 8 inches below equilibrium and released from rest. Find its
displacement for $t>0$.
\end{problem}

\begin{problem}\label{exer:6.2.8}
  A mass of 20 gm stretches a  spring 5 cm.  The spring
is attached to a dashpot with damping constant 400 dyne sec/cm.  Determine
the displacement for $t>0$ if the mass is  initially displaced
 9 cm  above equilibrium  and released from rest.

 \begin{solution}
     Since $k=\frac{mg}{\Delta l}=\frac{20\cdot980}{5}=3920$ the
equation of motion is $20y''+400y'+3920y=0$, or (A) $y''+20y'+196y=0$.
The characteristic polynomial of (A) is
$p(r)=r^2+20r+196=(r+10)^2+96$. Therefore,the general solution of (A)
is $y=e^{-10t}(c_1\cos4\sqrt6t+c_2\sin4\sqrt6t)$, so
$y'=-10y+4\sqrt6e^{-10t}(-c_1\sin4\sqrt6t+c_2\cos4\sqrt6t)$. Now
$y(0)=9$ and $y'(0)=0\Rightarrow c_1=9$ and $0=-90+4\sqrt6c_2$, or
$c_2=\frac{45}{2\sqrt6}$. Therefore,
$y=e^{-10t}\left(9\cos4\sqrt{6}t+\frac{45}{2\sqrt{6}}\sin4\sqrt{6}
t\right)$ cm.
 \end{solution}
\end{problem}

\begin{problem}\label{exer:6.2.9}
A 64 lb weight is suspended from a spring with constant $k=25$ lb/ft.
It is initially displaced 18 inches above equilibrium and released
from rest. Find its displacement for $t>0$ if the medium resists the
motion with 6 lb of force for each ft/sec of velocity.
\end{problem}

\begin{problem}\label{exer:6.2.10}
A 32 lb weight stretches a spring 1 ft in equilibrium. The weight is
initially displaced 6 inches above equilibrium and given a downward
velocity of 3 ft/sec. Find its displacement for $t>0$ if the medium
resists the motion with a force  equal to 3 times the speed
in ft/sec.

\begin{solution}
Since $k=\frac{mg}{\Delta l}=\frac{32}{1}=32$ the equation of
motion is (A) $y''+3y'+32y=0$. The characteristic polynomial of (A) is
$p(r)=r^2+3r+32=\left(r+\frac{3}{2}\right)^2+\frac{119}{4}$.
Therefore,the general solution of (A) is
$y=e^{-3t/2}\left(c_1\cos\frac{\sqrt{119}}{2}t
+c_2\sin\frac{\sqrt{119}}{2}t\right)$, so $y'=-\frac{3}{2}y+
\frac{\sqrt{119}}{2}e^{-3t/2}
\left(-c_1\sin\frac{\sqrt{119}}{2}t+c_2\cos\frac{\sqrt{119}}{2}t\right)$.
Now $y(0)=\frac{1}{2}$ and $y'(0)=-3\Rightarrow c_1=\frac{1}{2}$
and $-3=-\frac{3}{4}+\frac{\sqrt{119}}{2}c_2$, or
$c_2=-\frac{9}{2\sqrt{119}}$. Therefore,
$y=e^{-\frac{3}{2}t}\left(\frac{1}{2}\cos \frac{\sqrt{119}}{2} t
-\frac{9}{2\sqrt{119}}\sin\frac{\sqrt{119}}{2}t\right)$ ft.
\end{solution}
\end{problem}

\begin{problem}\label{exer:6.2.11}
An 8 lb weight stretches a spring 2 inches. It is attached to a
dashpot with damping constant $c$=4 lb-sec/ft. The weight is
initially displaced 3 inches above equilibrium and given a downward
velocity of 4 ft/sec. Find its displacement for $t>0$.
\end{problem}

\begin{problem}\label{exer:6.2.12}  
A 2 lb weight stretches a spring .32 ft. The weight is initially
displaced 4 inches below equilibrium and given an upward velocity of 5
ft/sec. The medium provides damping with constant $c=1/8$ lb-sec/ft.
Find and graph the displacement for $t>0$.

\begin{solution}
    Since $k=\frac{mg}{\Delta l}=\frac{2}{.32}=\frac{25}{4}$ the
equation of motion is
$\frac{1}{16}y''+\frac{1}{8}y'+\frac{25}{4}y=0$, or (A)
$y''+2y'+100y=0$. The characteristic polynomial of (A) is
$p(r)=r^2+2r+100=(r+1)^2+99$. Therefore,the general solution of (A) is
$y=e^{-t}(c_1\cos3\sqrt{11}t+c_2\sin3\sqrt{11}t)$, so
$y'=-y+3\sqrt{11}e^{-t}(-c_1\sin3\sqrt{11}t+c_2\cos3\sqrt{11}t)$. Now
$y(0)=-\frac{1}{3}$ and $y'(0)=5\Rightarrow c_1=-\frac{1}{3}$ and
$5=\frac{1}{3}+3\sqrt{11}c_2$, or $c_2=\frac{14}{9\sqrt{11}}$.
Therefore,$y=e^{-t}\left(-\frac{1}{3}\cos3\sqrt{11} t+
\frac{14}{9\sqrt{11}}\sin3\sqrt{11}t\right)$ ft.
\end{solution}
\end{problem}

\begin{problem}\label{exer:6.2.13}
  An 8 lb weight stretches a spring 8 inches in equilibrium.
It is attached to a dashpot with damping constant $c=.5$ lb-sec/ft
and subjected to an external force  $F(t)=4\cos2t$ lb.
Determine the steady state component of the displacement for $t>0$.
\end{problem}

\begin{problem}\label{exer:6.2.14}
A 32 lb weight stretches a spring 1 ft in equilibrium. It is attached
to a dashpot with constant $c=12$ lb-sec/ft. The weight is initially
displaced 8 inches above equilibrium and released from rest. Find its
displacement for $t>0$.

\begin{solution}
    Since $k=\frac{mg}{\Delta l}=32$ the
equation of motion is
(A) $y''+12y'+32y=0$.
The characteristic polynomial of (A) is
$p(r)=r^2+12r+32=(r+8)(r+4)$. Therefore,
the general solution of (A) is $y=c_1e^{-8t}+c_2e^{-4t}$, so
$y'=-8c_1e^{-8t}-4c_2e^{-4t}$. Now $y(0)=-\frac{2}{3}$ and
$y'(0)=0\Rightarrow c_1+c_2=\frac{2}{3},\-8c_1-4c_2=0$, so
$c_1=-\frac{2}{3}$,  $c_2=\frac{4}{3}$, and
$y=-\frac{2}{3}(e^{-8t}-2e^{-4t})$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:6.2.15}
A mass of one kg stretches a spring 49 cm in equilibrium. A dashpot
attached to the spring supplies a damping force of 4 N for each m/sec
of speed. The mass is initially displaced 10 cm above equilibrium and
given a downward velocity of 1 m/sec. Find its displacement for $t>0$.
\end{problem}

\begin{problem}\label{exer:6.2.16}
A mass of 100 grams stretches a spring 98 cm in equilibrium. A dashpot
attached to the spring supplies a damping force of 600 dynes for each
cm/sec of speed. The mass is initially displaced 10 cm above
equilibrium and given a downward velocity of 1 m/sec. Find its
displacement for $t>0$.

\begin{solution}
    Since $k=\frac{mg}{\Delta l}=\frac{100\cdot980}{98}=100$ the
equation of motion is
$100y''+600y'+1000y=0$, or
(A) $y''+6y'+10y=0$.
The characteristic polynomial of (A) is
$p(r)=r^2+6r+10=(r+3)^2+1$. Therefore,
the general solution of (A) is
$y=e^{-3t}(c_1\cos t+c_2\sin t)$, so
$y'=-3y+e^{-t}(-c_1\sin t+c_2\cos t)$.
Now $y(0)=10$ and $y'(0)=-100\Rightarrow c_1=10$ and
$-100=-30+c_2$, or $c_2=-70$. Therefore,
$y=e^{-3t}(10\cos t-70\sin t)$ cm.
\end{solution}
\end{problem}

\begin{problem}\label{exer:6.2.17}
A 192 lb weight is suspended from a spring with constant $k=6$ lb/ft
and subjected to an external force $F(t)=8\cos3t$ lb. Find the steady
state component of the displacement for $t>0$ if the medium resists
the motion with a force  equal to 8 times the speed in
ft/sec.
\end{problem}

\begin{problem}\label{exer:6.2.18}
A 2 gm mass is attached to a spring with constant 20 dyne/cm. Find the
steady state component of the displacement if the mass is subjected to
an external force $F(t)=3\cos4t-5\sin4t$ dynes and a dashpot supplies
4 dynes of damping for each cm/sec of velocity.

\begin{solution}
    The equation of motion is
(A) $2y''+4y'+20y=3\cos4t-5\sin4t$.
The steady  state component of the solution of (A)
is of the form
$y_p=A\cos4t+B\sin4t$; therefore $y_p'=-4A\sin4t+4B\cos4t$
and $y_p''=-16A\cos4t-16B\sin4t$, so
$2y_p''+4y_p'+20y_p=(-12A+16B)\cos4t-(16A+12B)\sin4t=3\cos4t-5\sin4t$
if $-12A+16B=3$, $-16A-12B=-5$; therefore $A=\frac{11}{100}$,
$B=\frac{27}{100}$, and
 $y_p=\frac{11}{100}\cos4t+\frac{27}{100}\sin4t$ cm.
\end{solution}
\end{problem}

\begin{problem}\label{exer:6.2.19}  
A 96 lb weight is attached to a spring with constant 12 lb/ft. Find
and graph the steady state component of the displacement if the mass is
subjected to an external force $F(t)=18\cos t-9\sin t$ lb and a
dashpot supplies 24 lb of damping for each ft/sec of velocity.
\end{problem}

\begin{problem}\label{exer:6.2.20}
A mass of one kg stretches a spring 49 cm in equilibrium. It is
attached to a dashpot that supplies a damping force of 4 N for each
m/sec of speed. Find the steady state component of its displacement if
it's subjected to an external force $F(t)=8\sin2t-6\cos2t$ N.

\begin{solution}
Since $k=\frac{mg}{\Delta l}=\frac{9.8}{.49}=20$ the
equation of motion is
(A) $y''+4y'+20y=8\sin2t-6\cos2t$.
The steady state component of the solution of (A)
is of the form
$y_p=A\cos2t+B\sin2t$; therefore $y_p'=-2A\sin2t+2B\cos2t$
and $y_p''=-4A\cos2t-4B\sin2t$, so
$y_p''+4y_p'+20y_p=(16A+8B)\cos2t-(8A-16B)\sin2t=8\sin2t-6\cos2t$
if $16A+8B=-6$, $-8A+16B=8$; therefore $A=-\frac{1}{2}$,
$B=\frac{1}{4}$, and
$y=-\frac{1}{2}\cos2t+\frac{1}{4}\sin2t$ m.
\end{solution}
\end{problem}

\begin{problem}\label{exer:6.2.21}
A mass $m$ is suspended from a spring with constant $k$ and subjected to an
external force $F(t)=\alpha\cos\omega_0t+\beta\sin\omega_0t$, where
$\omega_0$ is the natural frequency of the spring--mass system without
damping. Find the steady state component of the displacement if a dashpot
with constant $c$ supplies damping.
\end{problem}

\begin{problem}\label{exer:6.2.22}
Show that if $c_1$ and $c_2$ are not both zero then
$$
y=e^{r_1t}(c_1+c_2t)
$$
can't equal zero for more than one value of
$t$.

\begin{solution}
    If $e^{r_1t}(c_1+c_2t)=0$, then (A) $c_1+c_2t=0$. If $c_2=0$, then
$c_1\ne0$ (by assumption), so (A) is impossible. If $c_1\ne0$, then the
left side of (A) is strictly monotonic and therefore cannot have the
same value for two distinct values of $t$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:6.2.23}
Show that if $c_1$ and $c_2$ are not both zero then
$$
y=c_1e^{r_1t}+c_2e^{r_2t}
$$
 can't equal zero for more than one
value of $t$.
\end{problem}

\begin{problem}\label{exer:6.2.24}
Find the solution of the initial value problem
$$
my''+cy'+ky=0,\quad y(0)=y_0,\;y'(0)=v_0,
$$
given that the motion is underdamped, so  the general solution of
the equation is
$$
y=e^{-ct/2m}(c_1\cos\omega_1t+c_2\sin\omega_1t).
$$

\begin{solution}
    If $y=e^{-ct/2m}(c_1\cos\omega_1t+c_2\sin\omega_1t)$, then
$y'=-\frac{c}{2m}y+\omega_1
e^{-ct/2m}(-c_1\sin\omega_1t+c_2\cos\omega_1t)$, so $y(0)=y_0$ and
$y'(0)=v_0\Rightarrow c_1=y_0$ and
$v_0=-\frac{cy_0}{2m}+c_2\omega_1$, so $c_2=\frac{1}{\omega_1}
\left(v_0+\frac{cy_0}{2m}t\right)$ and $y=e^{-ct/2m}
\left(y_0\cos\omega_1t+\frac{1}{\omega_1}\left(v_0+\frac{cy_0}{
2m}\right)\sin\omega_1t\right)$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:6.2.25}
Find the solution of the initial value problem
$$
my''+cy'+ky=0,\quad y(0)=y_0,\;y'(0)=v_0,
$$
given that the motion is overdamped, so  the general solution of
the equation is
$$
y=c_1e^{r_1t}+c_2e^{r_2t}\;(r_1,r_2<0).
$$
\end{problem}

\begin{problem}\label{exer:6.2.26}
Find the solution of the initial value problem
$$
my''+cy'+ky=0,\quad y(0)=y_0,\;y'(0)=v_0,
$$
given that the motion is critically damped, so that the general
solution of the equation is of the form
$$
y=e^{r_1t}(c_1+c_2t)\,(r_1<0).
$$

\begin{solution}
    If $y=e^{r_1t}(c_1+c_2t)$, then $y'=r_1y+c_2e^{r_1t}$, so $y(0)=y_0$
and $y'(0)=v_0\Rightarrow c_1=y_0$ and $v_0=r_1y_0+c_2$, so
$c_2=v_0-r_1y_0$. Therefore,
$y=e^{r_1t}\left(y_0+(v_0-r_1y_0)t\right)$.
\end{solution}
\end{problem}

\end{document}