\documentclass{ximera}
\input{../preamble.tex}

\title{Exercises} \license{CC BY-NC-SA 4.0}

\begin{document}

\begin{abstract}
\end{abstract}
\maketitle

\begin{onlineOnly}
\section*{Exercises}
\end{onlineOnly}


\begin{problem}\label{exer:5.5.1} Find a particular solution. 

$y''+3y'+2y=7\cos x-\sin x$
\end{problem}

\begin{problem}\label{exer:5.5.2} Find a particular solution. 

$y''+3y'+y=(2-6x)\cos x-9\sin x$

\begin{solution}
Let
\begin{eqnarray*}
y_p&=&(A_0+A_1x)\cos x+(B_0+B_1x)\sin x; \text{ then}\\
y_p'&=&(A_1+B_0+B_1x)\cos x+(B_1-A_0-A_1x)\sin x\\
y_p''&=&(2B_1-A_0-A_1x)\cos x-(2A_1+B_0+B_1x)\sin x, \text{ so}
\end{eqnarray*}
\begin{eqnarray*}
y_p''+3y_p'+y_p&=&(3A_1+3B_0+2B_1+3B_1x)\cos x \\
&&+(3B_1-3A_0-2A_1-3A_1x)\sin x \\
&=&(2-6x)\cos x-9\sin x
\end{eqnarray*}
if $3B_1=-6$, $-3A_1=0$, $3B_0+3A_1+2B_1=2$, $-3A_0+3B_1+2A_1=-9$.
Therefore,$A_1=0$,   $B_1=-2$, $A_0=1$, $B_0=2$, and
 $y_p=\cos x+(2-2x)\sin x$.

\end{solution}

\end{problem}

\begin{problem}\label{exer:5.5.3} Find a particular solution. 

$y''+2y'+y=e^x(6\cos x+17\sin x)$
\end{problem}

\begin{problem}\label{exer:5.5.4} Find a particular solution. 

$y''+3y'-2y=-e^{2x}(5\cos2x+9\sin2x)$

\begin{solution}
Let $y=ue^{2x}$. Then
\begin{eqnarray*}
y''+3y'-2y&=&e^{2x}\left[(u''+4u'+4u)+3(u'+2u)-2u\right]\\
&=&e^{2x}(u''+7u'+8u)=-e^{2x}(5\cos2x+9\sin2x)
\end{eqnarray*}
if $u''+7u'+8u=-5\cos2x-9\sin2x$. Now let $u_p=A\cos2x+B\sin2x$. Then
\begin{eqnarray*}
u_p''+7u_p'+8u_p&=&-4(A\cos2x+B\sin2x)+14(-A\sin2x+B\cos2x)\\&&
+8(A\cos2x+B\sin2x)\\ &=&
(4A+14B)\cos2x-(14A-4B)\sin2x\\ &=&-5\cos2x-9\sin2x
\end{eqnarray*}
if $4A+14B=-5$, $-14A+4B=-9$. Therefore,$A=\frac{1}{2}$,
$B=-\frac{1}{2}$,
$$
 u_p=\frac{1}{2}(\cos2x-\sin2x), \text{ and }
y_p=\frac{{e^{2x}}{2}(\cos2x-\sin2x)}.
$$
\end{solution}
\end{problem}

\begin{problem}\label{exer:5.5.5} Find a particular solution. 

$y''-y'+y=e^x(2+x)\sin x$
\end{problem}

\begin{problem}\label{exer:5.5.6} Find a particular solution. 

$y''+3y'-2y=e^{-2x}\left[(4+20x)\cos 3x+(26-32x)\sin
3x\right]$

\begin{solution}
Let $y=ue^{-2x}$. Then
\begin{eqnarray*}
y''+3y'-2y&=&e^{-2x}\left[(u''-4u'+4u)+3(u'-2u)-2u\right]\\
&=&e^{-2x}(u''-u'-4u)\\ &=&
e^{-2x}\left[(4+20x)\cos 3x+(26-32x)\sin 3x\right]
\end{eqnarray*}
if $u''-u'-4u=(4+20x)\cos 3x+(26-32x)\sin 3x$.
Let
\begin{eqnarray*}
u_p&=&(A_0+A_1x)\cos3x+(B_0+B_1x)\sin3x; \text{ then}\\
u_p'&=&(A_1+3B_0+3B_1x)\cos3x+(B_1-3A_0-3A_1x)\sin3x\\
u_p''&=&(6B_1-9A_0-9A_1x)\cos3x-(2A_1+9B_0+9B_1x)\sin3x,\text{ so}\\
\end{eqnarray*}
\begin{eqnarray*}
u_p''-u_p'-4u_p&=&-\left[13A_0+A_1+3B_0-6B_1+(13A_1+3B_1)x\right]\cos3x
\\ &&-\left[13B_0+B_1-3A_0+6A_1+(13B_1-3A_1)x\right]\sin3x\\
&=&(4+20x)\cos 3x+(26-32x)\sin 3x \text{\ if}
\end{eqnarray*}
$$
\begin{array}{rcr}
-13A_1-13B_1&=&20\\ -13A_1-13B_1&=&-32
\end{array}
\text{ and }
\begin{array}{rcr}
-13A_0-13B_0-6A_1+6B_1&=&4.\\
-13A_0-13B_0-6A_1-6B_1&=&26.
\end{array}
$$
From the first two equations, $A_1=-2$, $B_1=2$. Substituting these in
the last two equations yields $-13A_0-3B_0=-10$, $3A_0-13B_0=16$.
Solving this pair yields $A_0=1$, $B_0=-1$. Therefore,
$$
 u_p=(1-2x)(\cos 3x-\sin 3x) \text{ and }
y_p=e^{-2x}(1-2x)(\cos 3x-\sin 3x).
$$
\end{solution}
\end{problem}

\begin{problem}\label{exer:5.5.7} Find a particular solution. 

$y''+4y=-12\cos2x-4\sin2x$
\end{problem}

\begin{problem}\label{exer:5.5.8} Find a particular solution. 

$y''+y=(-4+8x)\cos x+(8-4x)\sin x$

\begin{solution}
Let
\begin{eqnarray*}
y_p&=&(A_0x+A_1x^2)\cos x+(B_0x+B_1x^2)\sin x;\text{ then}\\
y_p'&=&\left[A_0+(2A_1+B_0)x+B_1x^2\right]\cos x
+\left[B_0+(2B_1-A_0)x-B_1x^2\right]\sin x\\
y_p''&=&
\left[2A_1+2B_0-(A_0-4B_1)x-A_1x^2\right]\cos x\\ &&+
\left[2B_1-2A_0-(B_0+4A_1)x-B_1x^2\right]\sin x, \text{ so}
\end{eqnarray*}
\begin{eqnarray*}
y_p''+y_p&=&(2A_1+2B_0+4B_1x)\cos x+(2B_1-2A_0-4A_1x)\sin x\\ &=&
(-4+8x)\cos x+(8-4x)\sin x
\end{eqnarray*}
if $4B_1=8$, $-4A_1=-4$, $2B_0+2A_1=-4$, $-2A_0+2B_1=8$. Therefore,
$A_1=1$, $B_1=2$, $A_0=-2$, $B_0=-3$, and $y_p=-x\left[(2-x)\cos
x+(3-2x)\sin x\right]$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:5.5.9} Find a particular solution. 

$4y''+y=-4\cos x/2-8x\sin x/2$
\end{problem}

\begin{problem}\label{exer:5.5.10} Find a particular solution. 

$y''+2y'+2y=e^{-x}(8\cos x-6\sin x)$

\begin{solution}
Let $y=ue^{-x}$. Then
\begin{eqnarray*}
y''+2y'+2y&=&e^{-x}\left[(u''-2u'+u)+2(u'-u)+2u\right]\\
&=&e^{-x}(u''+u)= e^{-x}(8\cos x-6\sin x)
\end{eqnarray*}
 if $u''+u=8\cos x-6\sin x$.
Now let
\begin{eqnarray*}
u_p&=&Ax\cos x+Bx\sin x; \text{ then}\\
u_p'&=&(A+Bx)\cos x+(B-Ax)\sin x\\
u_p''&=&(2B-Ax)\cos x-(2A+Bx)\sin x, \text{ so}
\end{eqnarray*}
$$
u_p''+u_p=2B\cos x-2A\sin x=8\cos x-6\sin x
$$
if $2B=8$, $-2A=-6$. Therefore,$A=3$, $B=4$, $u_p=x(3\cos x+4\sin x)$,
and $y_p=xe^{-x}(3\cos x+4\sin x)$.

\end{solution}
\end{problem}

\begin{problem}\label{exer:5.5.11} Find a particular solution. 

$y''-2y'+5y=e^x\left[(6+8x)\cos
2x+(6-8x)\sin2x\right]$
\end{problem}

\begin{problem}\label{exer:5.5.12} Find a particular solution. 

$y''+2y'+y=8x^2\cos x-4x\sin x$

\begin{solution}
Let
\begin{eqnarray*}
y_p&=&(A_0+A_1x+A_2x^2)\cos x +(B_0+B_1x+B_2x^2)\sin x;\text{ then}\\
y_p'&=&\left[A_1+B_0+(2A_2+B_1)x+B_2x^2\right]\cos x\\ &&
+\left[B_1-A_0+(2B_2-A_1)x-A_2x^2\right]\sin x, \\
y_p''&=&\left[-A_0+2A_2+2B_1-(A_1-4B_2)x-A_2x^2\right]\cos x\\ &&+
\left[-B_0+2B_2-2A_1-(B_1+4A_2)x-B_2x^2\right]\sin x,\text{ so}
\end{eqnarray*}
\begin{eqnarray*}
y_p''+2y_p'+y_p&=&2\left[A_1+A_2+B_0+B_1+(2A_2+B_1+2B_2)x+
B_2x^2\right]\cos x\\ && +
2\left[B_1+B_2-A_0-A_1+(2B_2-A_1-2A_2)x-A_2x^2\right]\sin x\\
&=&8x^2\cos x-4x\sin x \text{ if}
\end{eqnarray*}

\begin{enumerate}
    \item\label{solneqn:5.5.12A} \begin{eqnarray*}
    2B_2 &=& 8\\-2A_2 &=&
    0
    \end{eqnarray*},

    \item\label{solneqn:5.5.12B} \begin{eqnarray*}
    2B_1+4A_2+4B_2 &=& 0 \\ -2A_1-4A_2+4B_2 &=& -4
    \end{eqnarray*},

    \item \label{solneqn:5.5.12C} \begin{eqnarray*}
    2B_0+2A_1+2B_1+2A_2 &=& 0
\\ -2A_0-2A_1+2B_1+2B_2 &=& 0
    \end{eqnarray*},
\end{enumerate}


From \ref{solneqn:5.5.12A}, $A_2=0$, $B_2=4$. Substituting these into \ref{solneqn:5.5.12B} and solving
for $A_1$ and $B_1$ yields $A_1=10$, $B_1=-8$. Substituting the known
coefficients into \ref{solneqn:5.5.12C} and solving for $A_0$ and $B_0$ yields
$A_0=-14$, $B_0=-2$. Therefore,$y_p=-(14-10x)\cos x-(2+8x-4x^2)\sin
x$.

\end{solution}
\end{problem}

\begin{problem}\label{exer:5.5.13}   Find a particular solution. $y''+3y'+2y=(12+20x+10x^2)\cos x+8x\sin x$
\end{problem}

\begin{problem}\label{exer:5.5.14} Find a particular solution. $y''+3y'+2y=(1-x-4x^2)\cos2x-(1+7x+2x^2)\sin2x$
\end{problem}

\begin{problem}\label{exer:5.5.15}
Find a particular solution. $y''-5y'+6y=-e^x\left[(4+6x-x^2)\cos x-(2-4x+3x^2)\sin x\right]$
\end{problem}

\begin{problem}\label{exer:5.5.16}
Find a particular solution. $y''-2y'+y=-e^x\left[(3+4x-x^2)\cos x+(3-4x-x^2)\sin x\right]$
\end{problem}

\begin{problem}\label{exer:5.5.17}
Find a particular solution. $y''-2y'+2y=e^x\left[(2-2x-6x^2)\cos x+(2-10x+6x^2)\sin x\right]$
\end{problem}

\begin{problem}\label{exer:5.5.18} Find a particular solution and graph
it. $y''+2y'+y=e^{-x}\left[(5-2x)\cos
x-(3+3x)\sin x\right]$
\end{problem}

\begin{problem}\label{exer:5.5.19} Find a particular solution and graph
it.$y''+9y=-6\cos 3x-12\sin 3x$
\end{problem}

\begin{problem}\label{exer:5.5.20} Find a particular solution and graph
it. $y''+3y'+2y=(1-x-4x^2)\cos2x-(1+7x+2x^2)\sin2x$
\end{problem}

\begin{problem}\label{exer:5.5.21} Find a particular solution and graph
it.
$y''+4y'+3y=e^{-x}\left[(2+x+x^2)\cos x+(5+4x+2x^2)\sin x\right]$
\end{problem}

\begin{problem}\label{exer:5.5.22}
 Solve the
initial value problem. $y''-7y'+6y=-e^x(17\cos x-7\sin x), \quad   y(0)=4,\;  y'(0)=2$
\end{problem}

\begin{problem}\label{exer:5.5.23}
Solve the
initial value problem. $y''-2y'+2y=-e^x(6\cos x+4\sin x), \quad   y(0)=1,\;  y'(0)=4$
\end{problem}

\begin{problem}\label{exer:5.5.24} Solve the
initial value problem. $y''+6y'+10y=-40e^x\sin x, \quad  y(0)=2,\quad y'(0)=-3$
\end{problem}

\begin{problem}\label{exer:5.5.25} Solve the
initial value problem. $y''-6y'+10y=-e^{3x}(6\cos x+4\sin
x), \quad  y(0)=2,\quad y'(0)=7$
\end{problem}

\begin{problem}\label{exer:5.5.26}
Solve the
initial value problem. $y''-3y'+2y=e^{3x}\left[21\cos x-(11+10x)\sin x\right],
\;  y(0)=0, \quad   y'(0)=6$
\end{problem}

\begin{problem}\label{exer:5.5.27} Use the
principle
of superposition to find a particular solution. $y''-2y'-3y=4e^{3x}+e^x(\cos x-2\sin x)$
\end{problem}

\begin{problem}\label{exer:5.5.28} Use the
principle
of superposition to find a particular solution. $y''+y=4\cos x-2\sin x+xe^x+e^{-x}$
\end{problem}

\begin{problem}\label{exer:5.5.29} Use the
principle
of superposition to find a particular solution. $y''-3y'+2y=xe^x+2e^{2x}+\sin x$
\end{problem}

\begin{problem}\label{exer:5.5.30} Use the
principle
of superposition to find a particular solution. $y''-2y'+2y=4xe^x\cos x+xe^{-x}+1+x^2$
\end{problem}

\begin{problem}\label{exer:5.5.31} Use the
principle
of superposition to find a particular solution. $y''-4y'+4y=e^{2x}(1+x)+e^{2x}(\cos x-\sin
x)+3e^{3x}+1+x$
\end{problem}

\begin{problem}\label{exer:5.5.32}
Use the
principle
of superposition to find a particular solution, and
solve the initial value problem. $y''-4y'+4y=6e^{2x}+25\sin x , \quad   y(0)=5,\;  y'(0)=3$
\end{problem}

\begin{problem}\label{exer:5.5.33}
Solve the initial value problem and graph the solution. $y''+4y=-e^{-2x}\left[(4-7x)\cos x+(2-4x)\sin x\right],
\;  y(0)=3, \quad   y'(0)=1$
\end{problem}

\begin{problem}\label{exer:5.5.34} 
Solve the initial value problem and graph the solution. $y''+4y'+4y=2\cos2x+3\sin2x+e^{-x} , \quad   y(0)=-1,\;  y'(0)=2$
\end{problem}

\begin{problem}\label{exer:5.5.35} 
Solve the initial value problem and graph the solution. $y''+4y=e^x(11+15x)+8\cos2x-12\sin2x , \quad   y(0)=3,\;  y'(0)=5$
\end{problem}

\begin{problem}\label{exer:5.5.36}
\begin{enumerate}
\item\label{exer:5.5.36A} % (a)
Verify that if
$$
y_p=A(x)\cos\omega x+B(x)\sin\omega x
$$
where $A$ and $B$ are twice differentiable, then
\begin{eqnarray*}
 y_p'&=&(A'+\omega B)\cos\omega x+(B'-\omega A)
\sin\omega x \text{ and}\\
 y_p''&=&(A''+2\omega B'-\omega^2A)\cos\omega x
+(B''-2\omega A'-\omega^2B)\sin\omega x.
\end{eqnarray*}
\item\label{exer:5.5.36B} % (b)
Use the results of \ref{exer:5.5.36A} to verify that
\begin{eqnarray*}
ay_p''+by_p'+cy_p&=&\left[(c-a\omega^2)A+b\omega B+2a\omega
B'+bA'+aA''\right] \cos\omega x+\\  &&
\left[-b\omega A+(c-a\omega^2)B-2a\omega A'+bB'+aB''\right]\sin\omega
x.
\end{eqnarray*}
\item\label{exer:5.5.36C} % (c)
Use the results of \ref{exer:5.5.36A} to verify that
$$
y_p''+\omega^2 y_p=(A''+2\omega B')\cos\omega x+
(B''-2\omega A')\sin\omega x.
$$
\item % (d)
Prove Theorem~\ref{thmtype:5.5.2}.
\end{enumerate}
\end{problem}

\begin{problem}\label{exer:5.5.37}
Let $a$, $b$, $c$, and $\omega$ be constants, with $a\ne0$ and
$\omega>0$, and let
$$
P(x)=p_0+p_1x+\cdots+p_kx^k \quad \text{ and }\quad
Q(x)=q_0+q_1x+\cdots+q_kx^k,
$$
where at least one of the coefficients $p_k$, $q_k$ is nonzero, so
 $k$  is the larger of the degrees of $P$ and~$Q$.
\begin{enumerate}
\item % (a)
Show that if $\cos\omega x$ and $\sin\omega x$
are not solutions of the complementary equation
$$
ay''+by'+cy=0,
$$
then  there are polynomials
\begin{equation}\label{eq:eqA5.5.37}
A(x)=A_0+A_1x+\cdots+A_kx^k \quad  \text{and}\quad
B(x)=B_0+B_1x+\cdots+B_kx^k
\end{equation}
such that
$$
\begin{array}{lcl}
\quad (c-a\omega^2)A+b\omega B+2a\omega
B'+bA'+aA''&=&P\phantom{.}\\ 
-b\omega A+(c-a\omega^2)B-2a\omega A'+bB'+aB''&=&Q,
\end{array}
$$
where $(A_k,B_k)$, $(A_{k-1},B_{k-1})$, \dots,$(A_0,B_0)$ can be computed
successively by solving the systems
$$
\begin{array}{lcl}
\phantom{-}(c-a\omega^2)A_k+b\omega B_k&=&p_k\phantom{.}\\ 
-b\omega A_k+(c-a\omega^2)B_k&=&q_k,
\end{array}
$$
and, if $1\le r\le k$,
$$
\begin{array}{lcl}
\phantom{-}(c-a\omega^2)A_{k-r}+b\omega
B_{k-r}&=&p_{k-r}+\cdots\phantom{.}\\ 
-b\omega A_{k-r}+(c-a\omega^2)B_{k-r}&=&q_{k-r}+\cdots,
\end{array}
$$
where the terms indicated by ``$\cdots$'' depend upon the previously
computed coefficients with subscripts greater than $k-r$.
Conclude from this and Exercise~\ref{exer:5.5.36B} that
\begin{equation}\label{eq:eqB5.5.37}
y_p=A(x)\cos\omega x+B(x)\sin\omega x
\end{equation}
is a particular solution of
$$
ay''+by'+cy=P(x)\cos\omega x+Q(x)\sin\omega x.
$$
\item % (b)
Conclude from Exercise~\ref{exer:5.5.36C}
 that the equation
\begin{equation}\label{eq:eqC5.5.37}
a(y''+\omega^2 y)=P(x)\cos\omega x+Q(x)\sin\omega x
\end{equation}
does not have a solution of the form \ref{eq:eqB5.5.37} with  $A$
and $B$ as in \ref{eq:eqA5.5.37}. Then show that there are
polynomials
$$
A(x)=A_0x+A_1x^2+\cdots+A_kx^{k+1}\quad \text{ and } \quad
B(x)=B_0x+B_1x^2+\cdots+B_kx^{k+1}
$$
such that
$$
\begin{array}{rcl}
a(A''+2\omega B')&=&P\\ 
a(B''-2\omega A')&=&Q,
\end{array}
$$
where the pairs $(A_k,B_k)$, $(A_{k-1},B_{k-1})$, \dots, $(A_0,B_0)$ can be
computed successively as follows:
\begin{eqnarray*}
A_k&=&-\frac{q_k}{2a\omega(k+1)}\\ 
B_k&=&\phantom{-}\frac{p_k}{2a\omega(k+1)},
\end{eqnarray*}
and, if $k\ge 1$,
\begin{eqnarray*}
A_{k-j}&=&-\frac{1}{2\omega}
\left[\frac{q_{k-j}}{a(k-j+1)}-(k-j+2)B_{k-j+1}\right]\\ 
B_{k-j}&=&\phantom{-}\frac{1}{2\omega}
\left[\frac{p_{k-j}}{a(k-j+1)}-(k-j+2)A_{k-j+1}\right]
\end{eqnarray*}
for $1\le j\le k$.
Conclude that \ref{eq:eqB5.5.37} with this choice of the
polynomials $A$ and $B$ is a particular solution of
\ref{eq:eqC5.5.37}.
\end{enumerate}
\end{problem}

\begin{problem}\label{exer:5.5.38}
Show that  Theorem~\ref{thmtype:5.5.1} implies the following theorem:

Suppose  $\omega$ is a positive  number and $P$ and $Q$
are polynomials. Let $k$ be the larger of the degrees of $P$ and $Q$.
Then the equation
$$
ay''+by'+cy=e^{\lambda x}\left(P(x)\cos \omega x+Q(x)\sin \omega
x\right)
$$
has a particular solution
\begin{equation}\label{eq:eqA5.5.38}
y_p=e^{\lambda x}\left(A(x)\cos\omega x+B(x)\sin\omega x\right),
\end{equation}
where
$$
A(x)=A_0+A_1x+\cdots+A_kx^k \quad \text{ and }\quad
B(x)=B_0+B_1x+\cdots+B_kx^k,
$$
provided that $e^{\lambda x}\cos\omega x$ and $e^{\lambda x}\sin\omega
x$ are not solutions of  the complementary equation. The equation
$$
a\left[y''-2\lambda
y'+(\lambda^2+\omega^2)y\right]=
e^{\lambda x}\left(P(x)\cos \omega x+Q(x)\sin \omega x\right)
$$
$($for which $e^{\lambda x}\cos\omega x$ and $e^{\lambda x}\sin\omega
x$ are
solutions of the complementary equation$)$ has a particular solution
of the form \ref{eq:eqA5.5.38}, where
$$
A(x)=A_0x+A_1x^2+\cdots+A_kx^{k+1} \quad \text{ and }\quad
B(x)=B_0x+B_1x^2+\cdots+B_kx^{k+1}.
$$
\end{problem}

\begin{problem}\label{exer:5.5.39}
This exercise presents a method for evaluating the integral
$$
y=\int e^{\lambda x}\left(P(x)\cos \omega x+Q(x)\sin \omega x\right)\,dx
$$
where $\omega\ne0$ and
$$
P(x)=p_0+p_1x+\cdots+p_kx^k,\quad Q(x)=q_0+q_1x+\cdots+q_kx^k.
$$
\begin{enumerate}
\item % (a)
Show that $y=e^{\lambda x}u$, where
\begin{equation}\label{eq:eqA5.5.39}
u'+\lambda u=P(x)\cos \omega x+Q(x)\sin \omega x.
\end{equation}
\item % (b)
Show that \ref{eq:eqA5.5.39} has a particular solution of the form
$$
u_p=A(x)\cos \omega x+B(x)\sin \omega x,
$$
where
$$
A(x)=A_0+A_1x+\cdots+A_kx^k,\quad B(x)=B_0+B_1x+\cdots+B_kx^k,
$$
and  the pairs of coefficients
$(A_k,B_k)$, $(A_{k-1},B_{k-1})$, \dots,$(A_0,B_0)$
can be computed successively as the solutions of pairs of equations
obtained  by equating the coefficients
of $x^r\cos\omega x$ and $x^r\sin\omega x$ for $r=k$, $k-1$, \dots, $0$.
\item % (c)
Conclude that
$$
\int e^{\lambda x}\left(P(x)\cos \omega x+Q(x)\sin \omega x\right)\,dx
= e^{\lambda x}\left(A(x)\cos \omega x+B(x)\sin \omega x\right) +c,
$$
where $c$ is a constant of integration.
\end{enumerate}
\end{problem}

\begin{problem}\label{exer:5.5.40}
Use the method of Exercise~\ref{exer:5.5.39} to evaluate the integral.

\begin{enumerate}
    
\item  $\int x^2\cos x\,dx$ 

\item  $\int x^2e^x\cos x\,dx$ 

\item   \part{c} $\int xe^{-x}\sin2x\,dx$ 

\item \part{d} $\int x^2e^{-x}\sin x\,dx$ 

\item \part{e} $\int x^3e^x\sin x\,dx$ 

\item \part{f} $\int e^x\left[x\cos x-(1+3x)\sin x\right]\,dx$ 

\item \part{g} $\int e^{-x}\left[(1+x^2)\cos x+(1-x^2)\sin x\right]\,dx$
\end{enumerate}
\end{problem}

\end{document}