\documentclass{ximera}
\input{../preamble.tex}

\title{Exercises} \license{CC BY-NC-SA 4.0}

\begin{document}

\begin{abstract}
\end{abstract}
\maketitle

\begin{onlineOnly}
\section*{Exercises}
\end{onlineOnly}

\begin{problem}\label{exer:2.4.15} Solve the equation explicitly.
$$y'=\frac{y+x}{x}$$
\end{problem}

\begin{problem}\label{exer:2.4.16} Solve the equation explicitly.
$$y'=\frac{y^2+2xy}{x^2}$$



\begin{solution}
    $y=ux$;\ $u'x+u=u^2+2u$;\;
(A) $u'x=u(u+1)$. Since $u\equiv0$ and $u\equiv-1$ are constant
solutions of (A), $y\equiv0$ and $y=-x$ are
solutions of the given equation. The nonconstant solutions of (A)
satisfy
$=\frac{u'}{ u(u+1)}=\frac{1}{ x}$;\;
$\left[\frac{1}{ u}-\frac{1}{ u+1}\right]u'=\frac{1}{ x}$;\;
$\ln\left|\frac{u}{ u+1}\right|=\ln|x|+k$;\;
$\frac{u}{ u+1}=cx$;\ $u=(u+1)cx$;\ $u(1-cx)=cx$;\;
$u=\frac{cx}{1-cx}$;\;
$y=\frac{cx^2}{1-cx}$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.17} Solve the equation explicitly.
$$xy^3y'=y^4+x^4$$ 
\end{problem}

 \begin{problem}\label{exer:2.4.18} Solve the equation explicitly.
 $$y'=\frac{y}{x}+\sec\frac{y}{x}$$

 

 \begin{solution}
     $y=ux$;\ $u'x+u=u+\sec$;\;
$u'x=\sec u$;\;
$(\cos u)u'=\frac{1}{ x}$;\;
$\sin u=\ln|x|+c$;\;
$u=\sin^{-1}(\ln|x|+c)$;\;
$y=x\sin^{-1}(\ln|x|+c)$.
 \end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.19} Solve the equation explicitly.  Also, plot a direction field and some integral curves on the indicated rectangular region.
$$x^2y'=xy+x^2+y^2;   \quad \{-8\leq x\leq 8,-8\leq y\leq 8\}$$
\end{problem}

\begin{problem}\label{exer:2.4.20} Solve the equation explicitly.  Also, plot a direction field and some integral curves on the indicated rectangular region.
$$xyy'=x^2+2y^2;  \quad \{-4\leq x\leq 4,-4\leq y\leq 4\}$$



\begin{solution}
    Rewrite the given equation as $y'=\frac{x^2+2y^2}{ xy}$;\;
$y=ux$;\ $u'x+u=\frac{1}{ u}+2u$;\;
$u'x=\frac{1+u^2}{ u}$;\;
$\frac{uu'}{1+u^2}=\frac{1}{ x}$;\;
$\frac{1}{2}\ln(1+u^2)=\ln|x|+k$;\;
$\ln\left(1+\frac{y^2}{ x^2}\right)=\ln x^2+2k$;\;
$1+\frac{y^2}{ x^2}=cx^2$;\;
$x^2+y^2=cx^4$;\ $y=\pm x\sqrt{cx^2-1}$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.21} Solve the equation explicitly.  Also, plot a direction field and some integral curves on the indicated rectangular region.
$$y'=\frac{2y^2+x^2e^{-(y/x)^2}}{2xy};
\quad  \{-8\leq x\leq 8,-8\leq y\leq 8\}$$
\end{problem}

\begin{problem}\label{exer:2.4.22} Solve the initial value problem.
$$y'=\frac{xy+y^2}{x^2},
\quad y(-1)=2$$



\begin{solution}
    $y=ux$;\ $u'x+u=u+u^2$;\;
$u'x=u^2$;\;
$\frac{u'}{ u^2}=\frac{1}{ x}$;\;
$-\frac{1}{ u}=\ln|x|+c$;\;
$y(-1)=2\Rightarrow u(-1)=-2\Rightarrow c=\frac{1}{2}$;\;
$u=-\frac{2}{2\ln|x|+1}$;\;
$y=-\frac{2x}{2\ln|x|+1}$.
\end{solution}
\end{problem}

 \begin{problem}\label{exer:2.4.23} Solve the initial value problem.
 $$y'=\frac{x^3+y^3}{xy^2}, \quad y(1)=3$$
 \end{problem}

 \begin{problem}\label{exer:2.4.24} Solve the initial value problem.
 $$xyy'+x^2+y^2=0, \quad y(1)=2$$

 

 \begin{solution}
     Rewrite the given equation as $y'=-\frac{x^2+y^2}{ xy}$;\;
$y=ux$;\ $u'x+u=-\frac{1}{ u}-u$;\;
$u'x=-\frac{1+2u^2}{ u}$;\;
$-\frac{uu'}{1+2u^2}=\frac{1}{ x}$;\;
$-\frac{1}{4}\ln(1+2u^2)=\ln|x|+k$;\;
$x^4(1+2u^2)=c$;\;
$y(1)=2\Rightarrow u(1)=2\Rightarrow c=9$;\;
$x^4(1+2u^2)=9$;\;
$u^2=\frac{9-x^4}{2x^4}$;\;
$u=\frac{1}{ x^2}\left(\frac{9-x^4}{2}\right)^{1/2}$;\;
$y=\frac{1}{ x}\left(\frac{9-x^4}{2}\right)^{1/2}$.
 \end{solution}
 \end{problem}

\begin{problem}\label{exer:2.4.25} Solve the initial value problem.
$$y'=\frac{y^2-3xy-5x^2}{x^2}, \quad y(1)=-1$$
\end{problem}

\begin{problem}\label{exer:2.4.26}Solve the initial value problem.
$$x^2y'=2x^2+y^2+4xy, \quad y(1)=1$$



\begin{solution}
    Rewrite the given equation as $y'=2+\frac{y^2}{
x^2}+4\frac{y}{ x}$;\;
$y=ux$;\ $u'x+u=2+u^2+4u$;\;
$u'x=u^2+3u+2=(u+1)(u+2)$;\;
$\frac{u'}{(u+1)(u+2)}=\frac{1}{ x}$;\;
$\left[\frac{1}{ u+1}-\frac{1}{ u+2}\right]u'=\frac{1}{ x}$;\;
$\ln\left|\frac{u+1}{ u+2}\right|=\ln|x|+k$;\;
$\frac{u+1}{ u+2}=cx$;\;
$y(1)=1\Rightarrow u(1)=1\Rightarrow c=\frac{2}{3}$;\;
$\frac{u+1}{ u+2}=\frac{2}{3}x$;\;
$u+1=\frac{2}{3}x(u+2)$;\;
$u=\left(1-\frac{2}{3}x\right)=-1+\frac{4}{3}x$;\;
$u=-\frac{4x-3}{2x-3}$;\;
$y=-\frac{x(4x-3)}{2x-3}$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.27} Solve the initial value problem.
$$xyy'=3x^2+4y^2, \quad y(1)=\sqrt{3}$$
\end{problem}

\begin{problem}\label{exer:2.4.28} Solve the given homogeneous equation implicitly.
$$y'=\frac{x+y}{x-y}$$



\begin{solution}
    $y=ux$;\ $u'x+u=\frac{1+u}{1-u}$;\;
$u'x=\frac{1+u^2}{1-u}$;\;
$\frac{(1-u)u'}{1+u^2}=\frac{1}{ x}$;\;
$\tan^{-1}u-\frac{1}{2}\ln(1+u^2)=\ln|x|+c$;\;
$\tan^{-1}\frac{y}{
x}-\frac{1}{2}\ln\left(1+\frac{y^2}{ x^2}\right)=\ln|x|+c$;\;
$\tan^{-1}\frac{y}{
x}-\frac{1}{2}\ln(x^2+y^2)=c$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.29} Solve the given homogeneous equation implicitly.
$$(y'x-y)(\ln |y|-\ln |x|)=x$$
\end{problem}

\begin{problem}\label{exer:2.4.30} Solve the given homogeneous equation implicitly.
$$y'=\frac{y^3+2xy^2+x^2y+x^3}{x(y+x)^2}$$

Click here to see the answer.

\begin{solution}
    $y=ux$;
$u'x+u=\frac{u^3+2u^2+u+1}{(u+1)^2}=\frac{u(u+1)^2+1}{(u+1)^2}
=u+\frac{1}{(u+1)^2}$;
$u'x=\frac{1}{(u+1)^2}$;\;
$(u+1)^2u'=\frac{1}{ x}$;\;
$\frac{(u+1)^3}{3}=\ln|x|+c$;\;
$(u+1)^3=3(\ln|x|+c)$;\;
$\left(\frac{y}{ x}+1\right)^3=3(\ln|x|+c)$;\;
$(y+x)^3=3x^3(\ln|x|+c)$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.31}  Solve the given homogeneous equation implicitly.
$$y'=\frac{x+2y}{2x+y}$$
\end{problem}

\begin{problem}\label{exer:2.4.32}  Solve the given homogeneous equation implicitly.
$$y'=\frac{y}{y-2x}$$



\begin{solution}
    $y=ux$;\ $u'x+u=\frac{u}{ u-2}$;\;
(A)  $u'x=\frac{u(u-3)}{2-u}$;\;
Since $u\equiv0$ and $u\equiv3$
are constant solutions of (A), $y\equiv0$ and
$y=3x$ are solutions of the given equation.
The nonconstant solutions of (A) satisfy
$\frac{(2-u)u'}{ u(u-3)}=\frac{1}{ x}$;\;
$\left[\frac{1}{ u-3}+\frac{2}{
u}\right]u'=-\frac{3}{ x}$;\;
$\ln|u-3|+2\ln|u|=-3\ln|x|+k$;\;
$u^2(u-3)=\frac{c}{ x^3}$;\;
$y^2(y-3x)=c$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.33} Solve the given homogeneous equation implicitly.
$$y'=\frac{xy^2+2y^3}{x^3+x^2y+xy^2}$$
\end{problem}

\begin{problem}\label{exer:2.4.34} Solve the given homogeneous equation implicitly.
$$y'=\frac{x^3+x^2y+3y^3}{x^3+3xy^2}$$



\begin{solution}
    $y=ux$;\;
$u'x+u=
\frac {1+u+3u^3 }{ 1+3u^2}=u+{1 }{ 1+3u^2}$;\;
$(1+3u^2)u'=\frac{1}{ x}$;\;
$u+u^3=\ln |x|+c$;\;
$\frac{y}{ x}+\frac{y^3}{ x^3}=\ln |x|+c$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.35}
\begin{enumerate}
\item \label{partA:2.4.35} % (a)
Find a solution of the initial value problem
\begin{equation}\label{eqA:2.4.35}
x^2y'=y^2+xy-4x^2, \quad y(-1)=0
\end{equation}
on the interval $(-\infty,0)$. Verify that this solution is actually
valid on $(-\infty,\infty)$.
\item % (b)
Use Theorem~\ref{thmtype:2.3.1} to show that (\ref{eqA:2.4.35}) has a unique solution on $(-\infty,0)$.
\item % (c)
Plot a direction field for the differential equation in
(\ref{eqA:2.4.35}) on a square
$$
\{-r\leq x\leq r, -r\leq y\leq r\},
$$
where $r$
is any positive number. Graph the solution you obtained in
\ref{partA:2.4.35} on this field.
\item % (d)
Graph  other solutions  of (\ref{eqA:2.4.35}) that are defined on $(-\infty,\infty)$.
\item % (e)
Graph other solutions of (\ref{eqA:2.4.35})  that are defined only on intervals of
the form $(-\infty,a)$, where is a finite positive number.
\end{enumerate}
\end{problem}

\begin{problem}\label{exer:2.4.36}
\begin{enumerate}
\item % (a)
Solve the equation
\begin{equation}\label{eqA:2.4.36}
xyy'=x^2-xy+y^2
\end{equation}
implicitly.
\item % (b)
Plot a direction field for (\ref{eqA:2.4.36}) on a square
$$
\{0\leq x\leq r,0\leq y\leq r\}
$$
where $r$ is any positive number.
\item % (c)
Let $K$ be a positive integer. (You may have to try several
choices for $K$.)
Graph solutions of the initial value problems
$$
xyy'=x^2-xy+y^2,\quad y(r/2)=\frac{kr}{K},
$$
for $k=1$, $2$, \dots, $K$. Based on your observations, find
 conditions on the positive numbers $x_0$
and $y_0$ such that the initial value problem
\begin{equation}\label{eqB:2.4.36}
xyy'=x^2-xy+y^2,\quad y(x_0)=y_0,
\end{equation}
has  a unique solution (i) on $(0,\infty)$   or (ii) only
on an interval  $(a,\infty)$, where $a>0$?
\item % (d)
What can you say about the graph of the solution of (\ref{eqB:2.4.36})
as $x\to\infty$? (Again, assume that $x_0>0$ and $y_0>0$.)
\end{enumerate}



\begin{solution}
    Rewrite the given equation as $y'=\frac{x^2-xy+y^2}{ xy}$;\;
$y=ux$;\ $u'x+u=\frac{1}{ u}-1+u$;\;
$u'x=\frac{1-u}{ u}$;\;
$\frac{uu'}{ u-1}=-\frac{1}{ x}$;\;
$\left[1+\frac{1}{ u-1}\right]u'=-\frac{1}{ x}$;\;
$u+\ln|u-1|=-\ln|x|+k$;\;
$e^u(u-1)=\frac{c}{ x}$;\;
$e^{y/x}(y-x)=c$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.37}
\begin{enumerate}
\item % (a)
Solve the equation
\begin{equation}\label{eqA:2.4.37}
y'=\frac{2y^2-xy+2x^2}{xy+2x^2}
\end{equation}
implicitly.
\item % (b)
Plot a direction field for (\ref{eqA:2.4.37}) on a square
$$
\{-r\leq x\leq r,-r\leq y\leq r\}
$$
where $r$ is any positive number. By graphing solutions of (\ref{eqA:2.4.37}),
determine necessary and sufficient conditions on $(x_0,y_0)$ such that
(\ref{eqA:2.4.37}) has a solution on (i) $(-\infty,0)$ or (ii) $(0,\infty)$ such that
$y(x_0)=y_0$.
\end{enumerate}
\end{problem}

\begin{problem}\label{exer:2.4.38}
Follow the instructions of Exercise~\ref{exer:2.4.37} for the equation
$$
y'=\frac{xy+x^2+y^2}{xy}.
$$



\begin{solution}
    $y=ux$;\;
$u'x+u=1+\frac{1}{ u}+u$;\;
(A) $u'x=\frac{u+1}{ u}$.
Since (A)  has the constant solution $u=-1$;
$y=-x$ is
a solution of the given equation. The nonconstant solutions of (A)
satisfy
$\frac{uu'}{ u+1}=\frac{1}{ x}$;\;
$\left[1-\frac{1 }{ u+1}\right]u'=\frac{1}{ x}$;\;
$u-\ln |u+1|=\ln |x|+c$;\;
$\frac{y}{ x}-\ln\left|\frac{y}{ x}-1\right|=\ln|x|+c$;\;
$y-x\ln|y-x|=cy$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.39}
Pick any nonlinear homogeneous equation $y'=q(y/x)$  you like, and
plot  direction fields on the square $\{-r\leq x\leq r,\ -r\leq y\leq r\}$,
where $r>0$. What happens to the direction field as you vary $r$?
Why?
\end{problem}

\begin{problem}\label{exer:2.4.40}
Prove:  If $ad-bc\neq 0$, the equation
$$
y'=\frac{ax+by+\alpha}{cx+dy+\beta}
$$
can be transformed into the homogeneous nonlinear equation
$$
\frac{dY}{dX}=\frac{aX+bY}{cX+dY}
$$
by the substitution $x=X-X_0,\,  y=Y-Y_0$,
where $X_0$ and $Y_0$ are suitably chosen constants.

\begin{solution}
    If $x=X-X_0$ and $y=Y-Y_0$, then $\frac{dy}{ dx}=\frac{dY}{
dx}=\frac{dY}{ dX}\frac{dX}{ dx}=\frac{dY}{ dX}$, so $y=y(x)$
satisfies the given equation if and only if $Y=Y(X)$ satisfies
 $$
\frac{dY}{dX}=\frac{a(X-X_0)+b(Y-Y_0)+\alpha}{
c(X-X_0)+d(Y-Y_0)+\beta},
$$
 which reduces to the nonlinear homogeneous equation
$$
\frac{dY}{ dX}=\frac{aX+bY}{ cX+dY}
$$
 if and only if (B)

\begin{eqnarray*}
aX_0+bY_0&=&\alpha\\
cX_0+dY_0&=&\beta
\end{eqnarray*}

We will now show that
if $ad-bc\neq 0$, then it is possible (for any choice of $\alpha$
and $\beta$) to solve (B).
Multiplying the first equation in (B) by $d$ and
 the second by  $b$ yields
\begin{eqnarray*}
daX_0+dbY_0&=&d\alpha  \\
bcX_0+bdY_0&=&b\beta.
\end{eqnarray*}
Subtracting the second of these equations from the first yields
 $(ad-bc)X_0=\alpha d-\beta b$.
Since $ad-bc\neq 0$, this implies that $X_0=\frac{\alpha d-\beta b}{
ad-bc}$. Multiplying the first equation in (B) by
$c$ and the second by $a$ yields
\begin{eqnarray*}
caX_0+cbY_0&=&c\alpha  \\
acX_0+adY_0&=&a\beta.
\end{eqnarray*}
Subtracting the first of these equation from the second yields
 $(ad-bc)Y_0=\alpha c-\beta a$. Since $ad-bc\ne0$ this implies that
$Y_0=\frac{\alpha c-\beta a}{ ad-bc}$.
\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.41} Use a method suggested by Exercise~\ref{exer:2.4.40} to solve the
given equation implicitly.
$$y'=\frac{-6x+y-3}{2x-y-1}$$
\end{problem}

\begin{problem}\label{exer:2.4.42} Use a method suggested by Exercise~\ref{exer:2.4.40} to solve the
given equation implicitly.
$$y'=\frac{2x+y+1}{x+2y-4}$$



\begin{solution}
    For the given equation, (B)
of Exercise~\ref{exer:2.4.40} is
\begin{eqnarray*}
2X_0+Y_0&=&\phantom{-}1\\
X_0+2Y_0&=&-4.
\end{eqnarray*}
Solving this pair of equations yields   $X_0=2$ and $Y_0=-3$.
The transformed differential equation is (A)
$$
\frac{dY}{ dX}=\frac{2X+Y}{ X+2Y}.
$$
Let $Y=uX$;\;
$u'X+U=\frac{2+u}{1+2u}$;\;
(B) $u'X=-\frac{2(u-1)(u+1)}{ 2u+1}$. Since $u\equiv1$ and
$u\equiv-1$
satisfy (B), $Y=X$ and $Y=-X$ are solutions of (A).
 Since $X=x+2$ and $Y=y-3$, it follows that $y=x+5$
and $y=-x+1$ are solutions of the given equation. The nonconstant
solutions of (B) satisfy
$\frac{(2u+1)u'}{(u-1)(u+1)}=-\frac{2}{ X}$;\;
$\left[\frac{1}{ u+1}+\frac{3}{ u-1}\right]u'=-\frac{4}{
X}$;\;
$\ln|u+1|+3\ln|u-1|=-4\ln|X|+k$;\;
$(u+1)(u-1)^3=\frac{c}{ X^4}$;\;
$(Y+X)(Y-X)^3=c$;\;
Setting $X=x+2$ and $Y=y-3$ yields
$(y+x-1)(y-x-5)^3=c$.

\end{solution}
\end{problem}

\begin{problem}\label{exer:2.4.43} Use a method suggested by Exercise~\ref{exer:2.4.40} to solve the
given equation implicitly.
$$y'=\frac{-x+3y-14}{x+y-2}$$
\end{problem}

\begin{problem}\label{exer:2.4.54}
 Prove:  If $y$ is a solution of a homogeneous nonlinear equation
$y'=q(y/x)$,  so is $y_1=y(ax)/a$, where $a$ is any nonzero
constant.



\begin{solution}
    Differentiating  (A) $y_1(x)=\frac{y(ax)}{a}$   yields
(B) $y_1'(x)=\frac{1}{a}y'(ax)\cdot a=y'(ax)$. Since
$y'(x)=q(y(x)/x)$ on some interval $I$, (C) $y'(ax)=q(y(ax)/ax)$
on some interval $J$. Substituting (A) and (B) into (C)
yields $y_1'(x)=q(y_1(x)/x)$ on $J$.
\end{solution}
\end{problem}

\end{document}