\documentclass{ximera}
\input{../preamble.tex}

\title{Exercises} \license{CC BY-NC-SA 4.0}

\begin{document}

\begin{abstract}
\end{abstract}
\maketitle

\begin{onlineOnly}
\section*{Exercises}
\end{onlineOnly}



\begin{problem}\label{exer:8.3.1} Use the Laplace transform to solve the initial value problem. $y''+3y'+2y=e^t, \quad  y(0)=1,\quad y'(0)=-6$
\end{problem}

\begin{problem}\label{exer:8.3.2} Use the Laplace transform to solve the initial value problem. $y''-y'-6y=2, \quad  y(0)=1,\quad y'(0)=0$
\end{problem}

\begin{problem}\label{exer:8.3.3} Use the Laplace transform to solve the initial value problem. $y''+y'-2y=2e^{3t}, \quad  y(0)=-1,\quad y'(0)=4$
\end{problem}

\begin{problem}\label{exer:8.3.4} Use the Laplace transform to solve the initial value problem. $y''-4y=2 e^{3t}, \quad  y(0)=1,\quad y'(0)=-1$
\end{problem}

\begin{problem}\label{exer:8.3.5} Use the Laplace transform to solve the initial value problem. $y''+y'-2y=e^{3t}, \quad  y(0)=1,\quad y'(0)=-1$
\end{problem}

\begin{problem}\label{exer:8.3.6} Use the Laplace transform to solve the initial value problem. $y''+3y'+2y=6e^t, \quad  y(0)=1,\quad y'(0)=-1$
\end{problem}

\begin{problem}\label{exer:8.3.7} Use the Laplace transform to solve the initial value problem. $y''+y=\sin2t, \quad  y(0)=0,\quad y'(0)=1$
\end{problem}

\begin{problem}\label{exer:8.3.8} Use the Laplace transform to solve the initial value problem. $y''-3y'+2y=2e^{3t}, \quad  y(0)=1,\quad y'(0)=-1$
\end{problem}

\begin{problem}\label{exer:8.3.9} Use the Laplace transform to solve the initial value problem. $y''-3y'+2y=e^{4t}, \quad  y(0)=1,\quad y'(0)=-2$
\end{problem}

\begin{problem}\label{exer:8.3.10} Use the Laplace transform to solve the initial value problem. $y''-3y'+2y=e^{3t}, \quad  y(0)=-1,\quad y'(0)=-4$
\end{problem}

\begin{problem}\label{exer:8.3.11} Use the Laplace transform to solve the initial value problem. $y''+3y'+2y=2e^t, \quad  y(0)=0,\quad y'(0)=-1$
\end{problem}

\begin{problem}\label{exer:8.3.12} Use the Laplace transform to solve the initial value problem. $y''+y'-2y=-4, \quad  y(0)=2,\quad y'(0)=3$
\end{problem}

\begin{problem}\label{exer:8.3.13} Use the Laplace transform to solve the initial value problem. $y''+4y=4, \quad  y(0)=0,\quad y'(0)=1$
\end{problem}

\begin{problem}\label{exer:8.3.14} Use the Laplace transform to solve the initial value problem. $y''-y'-6y=2, \quad  y(0)=1,\quad y'(0)=0$
\end{problem}

\begin{problem}\label{exer:8.3.15} Use the Laplace transform to solve the initial value problem. $y''+3y'+2y=e^t, \quad  y(0)=0,\quad y'(0)=1$
\end{problem}

\begin{problem}\label{exer:8.3.16} Use the Laplace transform to solve the initial value problem. $y''-y=1, \quad  y(0)=1,\quad y'(0)=0$
\end{problem}

\begin{problem}\label{exer:8.3.17} Use the Laplace transform to solve the initial value problem. $y''+4y=3\sin t, \quad  y(0)=1,\quad y'(0)=-1$
\end{problem}

\begin{problem}\label{exer:8.3.18} Use the Laplace transform to solve the initial value problem. $y''+y'=2e^{3t}, \quad  y(0)=-1,\quad y'(0)=4$
\end{problem}

\begin{problem}\label{exer:8.3.19} Use the Laplace transform to solve the initial value problem. $y''+y=1, \quad  y(0)=2,\quad y'(0)=0$
\end{problem}

\begin{problem}\label{exer:8.3.20} Use the Laplace transform to solve the initial value problem. $y''+y=t, \quad  y(0)=0,\quad y'(0)=2$
\end{problem}

\begin{problem}\label{exer:8.3.21} Use the Laplace transform to solve the initial value problem. $y''+y=t-3\sin2t, \quad  y(0)=1,\quad y'(0)=-3$
\end{problem}

\begin{problem}\label{exer:8.3.22} Use the Laplace transform to solve the initial value problem. $y''+5y'+6y=2e^{-t}, \quad  y(0)=1,\quad y'(0)=3$
\end{problem}

\begin{problem}\label{exer:8.3.23} Use the Laplace transform to solve the initial value problem. $y''+2y'+y=6\sin t-4\cos t, \quad  y(0)=-1,\;
y'(0)=1$
\end{problem}

\begin{problem}\label{exer:8.3.24} Use the Laplace transform to solve the initial value problem. $y''-2y'-3y=10\cos t, \quad  y(0)=2,\quad y'(0)=7$
\end{problem}

\begin{problem}\label{exer:8.3.25} Use the Laplace transform to solve the initial value problem. $y''+y=4\sin t+6\cos t, \quad  y(0)=-6,\;
y'(0)=2$
\end{problem}

\begin{problem}\label{exer:8.3.26} Use the Laplace transform to solve the initial value problem.  $y''+4y=8\sin2t+9\cos t, \quad  y(0)=1,\quad y'(0)=0$
\end{problem}

\begin{problem}\label{exer:8.3.27} Use the Laplace transform to solve the initial value problem. $y''-5y'+6y=10e^t\cos t, \quad  y(0)=2,\quad y'(0)=1$
\end{problem}

\begin{problem}\label{exer:8.3.28} Use the Laplace transform to solve the initial value problem. $y''+2y'+2y=2t, \quad  y(0)=2,\quad y'(0)=-7$
\end{problem}

\begin{problem}\label{exer:8.3.29} Use the Laplace transform to solve the initial value problem. $y''-2y'+2y=5\sin t+10\cos t, \quad  y(0)=1,\;
y'(0)=2$
\end{problem}

\begin{problem}\label{exer:8.3.30} Use the Laplace transform to solve the initial value problem. $y''+4y'+13y=10e^{-t}-36e^t, \quad  y(0)=0,\;
y'(0)=-16$
\end{problem}

\begin{problem}\label{exer:8.3.31} Use the Laplace transform to solve the initial value problem. $y''+4y'+5y=e^{-t}(\cos t+3\sin t), \quad
y(0)=0,\quad y'(0)=4$
\end{problem}

\begin{problem}\label{exer:8.3.32} Use the Laplace transform to solve the initial value problem. $2y''-3y'-2y=4e^t, \quad   y(0)=1,\;  y'(0)=-2$
\end{problem}

\begin{problem}\label{exer:8.3.33} Use the Laplace transform to solve the initial value problem. $6y''-y'-y=3e^{2t}, \quad   y(0)=0,\;  y'(0)=0$
\end{problem}

\begin{problem}\label{exer:8.3.34} Use the Laplace transform to solve the initial value problem. $2y''+2y'+y=2t, \quad   y(0)=1,\;  y'(0)=-1$
\end{problem}

\begin{problem}\label{exer:8.3.35} Use the Laplace transform to solve the initial value problem. $4y''-4y'+5y=4\sin t-4\cos t, \quad   y(0)=0,\;  y'(0)=11/17$
\end{problem}

\begin{problem}\label{exer:8.3.36} Use the Laplace transform to solve the initial value problem. $4y''+4y'+y=3\sin t+\cos t, \quad   y(0)=2,\;  y'(0)=-1$
\end{problem}

\begin{problem}\label{exer:8.3.37} Use the Laplace transform to solve the initial value problem. $9y''+6y'+y=3e^{3t}, \quad   y(0)=0,\;  y'(0)=-3$
\end{problem}

\begin{problem}\label{exer:8.3.38}
Suppose $a,b$, and $c$ are constants and $a\ne0$.  Let
$$
y_1={\cal L}^{-1}\left(as+b\over as^2+bs+c\right)\mbox{\quad and \quad}
y_2={\cal L}^{-1}\left(a\over as^2+bs+c\right).
$$
Show that
$$
y_1(0)=1,\quad y_1'(0)=0\mbox{\quad and \quad} y_2(0)=0,\quad y_2'(0)=1.
$$
\begin{hint}
Use the Laplace transform to solve the initial value
problems
\begin{eqnarray*}
ay''+by'+cy&=&0, \quad   y(0)=1,\quad y'(0)=0 \\
ay''+by'+cy&=&0, \quad  y(0)=0,\quad y'(0)=1.
\end{eqnarray*}
\end{hint}
\end{problem}

\end{document}